Glial ferritin maintains neural stem cells via transporting iron required for self-renewal in Drosophila

Version 1

Abstract

Abstract

Stem cell niche is critical for regulating the behavior of stem cells. Drosophila neural stem cells (Neuroblasts, NBs) are encased by glial niche cells closely, but it still remains unclear whether glial niche cells can regulate the self-renewal and differentiation of NBs. Here we show that ferritin produced by glia, cooperates with Zip13 to transport iron into NBs for the energy production, which is essential to the self-renewal and proliferation of NBs. The knockdown of glial ferritin encoding genes causes energy shortage in NBs, which leads to the low proliferation and premature differentiation of NBs. Moreover, the level of glial ferritin production is affected by the status of NBs, establishing a bicellular iron homeostasis. In this study, we demonstrate that glial cells are indispensable to maintain the self-renewal of NBs, unveiling a novel role of the NB glial niche during brain development.

Introduction

Neural stem cells (NSCs) possess the remarkable ability to self-renew for maintaining their identity and differentiate into diverse neural cell types required for neurogenesis during brain development. The self-renewal, differentiation, and proliferation of NSCs are regulated by both intrinsic programs inside stem cells and extrinsic cues from the surrounding niches. The mammalian NSC niche found in the neurogenic regions is composed of diverse cell types including choroid plexus, vascular endothelia cells, pericytes, microglia, and meninges, which regulate NSC behaviors such as proliferation, self-renewal, and differentiation (Bjornsson, Apostolopoulou, Tian, & Temple, 2015). The diversity and complexity of these cell populations confer niche the ability of integrating local and systemic signals, but make it complicated to study the interaction between niches and NSCs in mammals, which is a critical issue to understand neurogenesis and improve clinical applications of NSCs.

Drosophila larval central nervous system (CNS), composed of central brain (CB), optic lobe (OL) and ventral nerve cord (VNC), is a simple but sophisticated model to decipher the regulatory mechanisms underlying NSC behaviors mediated by niche signals. In this system, Drosophila NSCs, called neuroblasts (NBs), and their progenies are encased by a variety of glial cells including perineurial glia, subperineurial glia, cortex glia, astrocyte-like glia and ensheathing glia (Hartenstein, 2011; Ito, Urban, & Technau, 1995; Pereanu, Shy, & Hartenstein, 2005). These glial cells not only form the physical support structure for NBs, but also provide extrinsic cues that regulate NB reactivation, survival, and proliferation (Read, 2018; Speder & Brand, 2014). However, it remains elusive whether the glial-cell-forming niche is involved in the regulation of the prominent characteristics of NBs, self-renewal and differentiation.

Iron is a vital micronutrient for almost all living organisms since it participates in many crucial biological processes, including mitochondrial electron transport and cellular respiration that both support ATP production used for most cellular activities. High levels of iron are found in many organs including the brain, where iron is required for myelination and formation of neuronal dendritic trees (Lozoff et al., 2006; Moller et al., 2019; Rice & Barone, 2000). Iron deficiency in infants and early childhood has been well documented to be associated with long-term cognitive, social, and emotional problems (Fretham, Carlson, & Georgieff, 2011). Thus, it is intriguing to investigate whether iron is also required in NSCs during neurogenesis.

Ferritin, a ubiquitous protein present in most organisms, is composed of 24 subunits that assemble into a hollow-sphere complex capable of storing up to 4,500 iron atoms in a bioavailable form (Andrews, 2005; Knovich, Storey, Coffman, Torti, & Torti, 2009). Ferritin in both mammal and insect is a heteropolymer, consisting of heavy and light chain homologue subunits. In Drosophila, these subunits are encoded by ferritin 1 heavy chain homologue (Fer1HCH) and ferritin 2 light chain homologue (Fer2LCH) respectively (Arosio, Ingrassia, & Cavadini, 2009; Pham & Winzerling, 2010), which both contain secretion signal peptides that direct them to the ER during translation (Lind, Ekengren, Melefors, & Soderhall, 1998). Consequently, Drosophila ferritin is confined to the secretory pathway, abundant in the hemolymph, and absent from the cytosol (Nichol, Law, & Winzerling, 2002). Previous researches have shown that Drosophila ferritin not only directly participates in dietary iron absorption by removing iron from enterocytes across the basolateral membrane, but competes with transferrin to deliver iron between the gut and the fat body (Tang & Zhou, 2013; Xiao, Liu, Zhao, Wang, & Zhou, 2019). These findings underscore the pivotal role of ferritin in iron transport. Interestingly, iron-loaded ferritin, serving as an essential mitogen, promotes proliferation of cultured Drosophila cells (Li, 2010). Therefore, it is tempting to investigate whether ferritin is required for self-renewal and proliferation of NSCs during brain development.

Here we show that glial ferritin supplies iron into NBs to form iron-sulfur (Fe-S) clusters, which is necessary for the production of ATP, providing the energy required for self-renewal and proliferation of NBs, highlighting ferritin as a potential target for suppressing tumor.

Results

Glial ferritin is required for NB maintenance and proliferation

To identify new niche factors involved in the balance between NB self-renewal and differentiation, we used repo-GAL4 to perform an RNAi mediated screen in glial cells. Fer1HCH and Fer2LCH, two ferritin subunit encoding genes, were identified to be essential for NB maintenance and brain development. Specifically, knocking down Fer1HCH or Fer2LCH in glial cells significantly reduced the number of NBs in the CB and VNC at the third-instar larval stage, which resulted in decreased size of the CNS (Figure 1A and B and Figure 1-figure supplement 1A). To further assess the impact on NB proliferation, we employed the mitotic marker phosphohistone H3 (PH3) and the thymidine analogue 5-ethynyl-2’-deoxyuridine (EdU), revealing a substantial decrease in the proliferation rate of NBs upon glial ferritin knockdown (Figure 1A and C and Figure 1-figure supplement 1B and C). These results indicated that glial ferritin is required for maintaining NBs and their proliferation.

Ferritin in glia is required for NB maintenance and proliferation

(A) NB number and proliferation rate after Fer1HCH or Fer2LCH RNAi in glia.

(B) Quantification of NB number in CB (A).

(C) Quantification of NB proliferation rate in CB (A).

(D and E) RT-PCR analysis of Fer1HCH and Fer2LCH mRNA levels (D) and normalized quantification (E) after glial ferritin knockdown.

(Fand H) Western blot of Fer1HCH (F) and Fer2LCH (H) after ferritin knockdown in glia. Tubulin was used as a loading control.

(G and I) Quantification of normalized Fer1HCH intensity in (G) and Fer2LCH intensity in (I).

(J) Rescue of NB number and proliferation when simultaneously overexpressing RNAi-resistant

Fer1HCH and Fer1HCH RNAi (or Fer2LCH and Fer2LCH RNAi) in glia. (K and L) Quantification of NB number and proliferation rate in (J).

(B, n=11; C, n=10; E, G and I, n=3; K, n=8,8,9,7,9; L, n=9,8,8,7,12; ns, not significant; *: p<0.05; **: p<0.01; ***: p<0.001; ****: p<0.0001; Dpn, NB nuclei, red; PH3, green; DAPI, nuclei, blue).

To identify the glial subpopulation in which ferritin functions, we employed subtype-restricted GAL4 to perform ferritin knockdowns (Table 1). We observed that knocking down ferritin genes in cortex glia led to a similar yet milder phenotype compared to that observed in the pan-glia knockdown (Figure 1-figure supplement 1D). This milder effect may be attributed to potential compensatory mechanisms from other glial subpopulations. Interestingly, knocking down ferritin in other glial subtypes did not yield any discernible phenotype. These findings suggest that ferritin functions collectively across all glial populations to regulate the activity of NBs.

We confirmed the phenotype with multiple independent RNAi lines for Fer1HCH or Fer2LCH, and observed similar abnormalities of larval NBs and brain development (Figure 1-figure supplement 2A). Hereafter, UAS-Fer1HCH-RNAi (THU5585) and UAS-Fer2LCH-RNAi (TH01861.N) were used for most studies, and we only quantified NB number and proliferation rate in the CB for convenience in this research.

To evaluate the efficiency of RNAi for Fer1HCH or Fer2LCH, RT-PCR was performed with total RNA isolated from whole brain dissected from third-instar larvae. The abundance of Fer1HCH or Fer2LCH mRNA was significantly reduced in glial Fer1HCH or Fer2LCH knockdown respectively (Figure 1D and E). Western blot analysis of larval brain protein extract using polyclonal antibodies against Fer1HCH or Fer2LCH further confirmed that both subunits were down-regulated in glial knockdown of either subunit (Figure 1F-I), consistent with previous research in Drosophila (Missirlis et al., 2007). To exclude the possibility of RNAi off-target effects, we used repo-GAL4 to overexpress RNAi-resistant Fer1HCH (or Fer2LCH) containing silent mutations in the region targeted by shRNA, which resulted in a significant increase in the number and proliferation rate of NBs compared to glial ferritin knockdown alone (Figure 1J-L and Figure 1-figure supplement 2B, C). Additionally, similar but weaker phenotypes were observed when Fer1HCHDN, a ferroxidase-inactive form of Fer1HCH (Missirlis et al., 2007), was overexpressed in glial cells (Figure 1-figure supplement 2D). These results support the conclusion that ferritin in glial cells is essential for the maintenance and proliferation of NBs, suggesting that iron homeostasis is involved in this process.

Ferritin is produced mainly in glial cells and secreted into NBs through a vesicle-dependent pathway

To investigate the role of iron homeostasis in Drosophila CNS, we tried to knock down genes involved in iron transport, storage, and regulation in NBs, neurons, and glial cells, respectively. Interestingly, only glial knockdown of ferritin-related genes led to NB loss and defective brain development. It appears that ferritin generated in glial cells is the only crucial protein for regulating the behavior of NBs and, ultimately, the larval brain development. To recapitulate the endogenous ferritin expression pattern, we took advantage of a Fer1HCH/Fer2LCH enhancer trapped GAL4 line to drive a variant of DsRed protein with a nuclear localization signal (DsRed Stinger). Surprisingly, our immunostaining experiments revealed that most, if not all, DsRed was co-localized with the glia nuclear protein Repo, instead of the NB nuclear protein Dpn (Figure 2A). These results further confirm that glial cells are the main producers of ferritin in the larval brain.

Ferritin is produced mainly in glial cells in the Drosophila CNS and secreted into NBs through a vesicle dependent pathway

(A) The pattern of Dsred Stinger driven by Fer1HCH/Fer2LCH-GAL4 in CNS.

(B) The distribution of ferritin labeled by Fer1HCHG188 in CNS of control and glial ferritin knockdown.

(C) Overexpressed ferritin tagged with GFP and mCherry in glia secreted into NBs.

(D) The distribution of Fer1HCHG188 when blocking glial vesicular trafficking via overexpressing dominant-negative dynamin.

(E) Quantification of (D), GFP signal in NBs when overexpressing shi ts in glia. (E, n=50, 48; **: p<0.01)

Fer1HCHG188 encodes a GFP-tagged Fer1HCH subunit (Missirlis et al., 2007), which labels the distribution of Fer1HCH, representing the pattern of ferritin. As shown in the Figure 2B, ferritin was detected in NBs besides glial cells, despite that ferritin is specifically produced in glial cells. To explore the possibility of ferritin transfer from glial cells to NBs, we observed the GFP signal of Fer1HCHG188 with glial ferritin knockdown, and found that GFP in the brain was almost completely depleted (Figure 2B). Furthermore, we used repo-GAL4 to drive the expression of GFP-tagged Fer1HCH and mCherry-tagged Fer2LCH in glial cells. We observed that both Fer1HCH::GFP and Fer2LCH::mCherry were present in NBs, while the control GFP and mCherry were restricted to glial cells (Figure 2C). Altogether, we conclude that ferritin is generated in glial cells and then transferred to NBs in the larval brain.

In Drosophila, both ferritin subunits contain secretion signal peptides that direct protein to the endoplasmic reticulum during translation, which allows the ferritin to be transported out of the cell. To investigate whether the ferritin was transported from glia to NBs via the vesicular transport, we expressed a temperature-sensitive dominant-negative Shibire dynamin protein (shits) to block vesicle trafficking from glial cells, and found that Fer1HCH::GFP accumulated in glia, accompanying with significantly diminished Fer1HCH signal in NBs (Figure 2D and E), suggesting glial ferritin secretes into NBs through vesicle system.

Glial ferritin defects lead to iron deficiency in NBs

Given that ferritin is required for iron storage as well as delivery, glial ferritin knockdown might increase free iron level, leading to glial iron overload. Alternatively, it will block the iron transport from glia to NBs, resulting in NB iron deficiency. To determine the underlying cause of the brain development defects observed upon glial ferritin knockdown, we used iron chelator or ferric ammonium citrate (FAC) to manipulate iron intake and detect how it will influence NBs. We observed that depriving iron intake by adding iron chelator bathophenanthrolinedisulfonic acid disodium (BPS) to the food led to a dramatic decrease in the number and proliferation rate of NBs, compared to control larvae on normal food (Figure 3A). In line with this result, blocking iron absorption by knocking down intestinal ferritin (Tang & Zhou, 2013) resulted in the similar defects of brain (Figure 3-figure supplement 1A-C). However, supplementing the diet with FAC did not induce any observable defects (Figure 3A). We also tested another iron chelator deferiprone (DFP), and found that DFP alone did not cause any brain defects in wild-type (WT) larvae, but it significantly exacerbated the effect of glial ferritin knockdown on defective NB maintenance (Figure 3B and C). Furthermore, we attempted to rescue the brain defects observed in glial ferritin knockdown by providing dietary FAC, but the effect was negligible (Figure 3-figure supplement 1D and E). This may be due to the possibility that the knockdown of glial ferritin largely, if not completely, prevented the transport of iron from glia to NBs, which hindered the additional iron from entering NBs. These results indicate that the NB loss phenotype induced by glial ferritin knockdown is primarily due to iron deficiency in NBs.

Glial ferritin defects lead to iron deficiency in NBs

(A) The conditions of brain development in Drosophila with feeding iron chelator BPS or supplementary iron FAC.

(B) Iron depletion by adding DFP to food exacerbated the phenotype of NB loss induced by glial ferritin knockdown.

(C) Quantification of NB number in (B).

(D) Zip13 in CNS is produced mainly by glial cells, but not NBs.

(E) Brain defects in Zip13 knockdown or double knockdown of Zip13 and Fer2LCH. (F and G) Quantification of NB number and proliferation rate in (E).

(H) Iron supplement can rescue the NB proliferation in Zip13 RNAi driven by repo-GAL4.

(I) Quantification of NB proliferation in (H).

(C, n=6; F, n=6,10,11,8; G, n=13,10,10,5; I, n=8; ns, not significant; *: p<0.05; **: p<0.01; ****: p<0.0001)

Zip13, generally known as a zinc transporter, functions as an iron transporter, cooperating with ferritin on iron transport in Drosophila intestine (Xiao, Wan, Fan, Tang, & Zhou, 2014). Specifically, knockdown of Zip13 leads to failure of iron loading in secretory vesicles and consequently defective iron efflux (Xiao et al., 2014). To investigate whether the NB loss phenotype of glial ferritin knockdown is caused by defective iron transport, we next sought to identify the role of Zip13 in the larval brain. Firstly, we examined the expression pattern of Zip13 and found that, similar to ferritin, the DsRed stinger driven by Zip13-GAL4 was co-localized with Repo in glial cells rather than Dpn in NBs (Figure 3D). Subsequently, we knocked down Zip13 in glial cells and observed a similar but weaker phenotype with the slightly reduced NB number and proliferation rate (Figure 3E-G). Consistently, the defective brain development was confirmed in Zip13 mutant (Figure 3-figure supplement 1F-H). Moreover, the effect of double knockdown of both Zip13 and Fer2LCH was much more severe than that of either single knockdown, with a significant reduction in NB number and proliferation rate (Figure 3E-G). Furthermore, we found that the reduced proliferation rate caused by Zip13 knockdown could be partially rescued by iron supplementation in the food (Figure 3H and I). Overall, these results suggest that glial ferritin collaborates with Zip13 to deliver iron from glia to NBs.

Since previous studies have demonstrated the essential role of glial cells in NB survival and proliferation (Bailey et al., 2015; Read, 2018), it is possible that iron overload in glial cells up-regulated the level of reactive oxygen species (ROS), leading to apoptosis of glial cells, which ultimately resulted in the loss of NBs. To rule out this possibility, firstly, we examined the ROS level of glial ferritin knockdown via a transgenic reporter GstD-GFP for oxidative stress signaling (Sykiotis & Bohmann, 2008), and found an obvious increase as expected (Figure 3-figure supplement 2A and B). However, overexpressing anti-oxidant genes in glial cells or NBs did not restore the NB number and proliferation (Figure 3-figure supplement 2C and D). ROS accumulation leads to lipid peroxidation, which in turn results in ferroptosis (Hirschhorn & Stockwell, 2019). Mitochondria in cells undergoing ferroptosis typically exhibit a characteristic shrinkage and damage, while the nuclei remain structurally intact (Mumbauer, Pascual, Kolotuev, & Hamaratoglu, 2019). In order to investigate whether the NB loss is due to ferroptosis, we examined the mitochondrial morphology using Transmission Electron Microscopy (TEM). However, our observation revealed no discernible defects in the mitochondria of NBs after glial ferritin knockdown (Figure 3-figure supplement 2E and F). Taken with the findings from iron chelator feeding together, we conclude that NB loss resulting from glial ferritin knockdown is not a consequence of ferroptosis. Secondly, we examined the apoptosis in glial cells via Caspase-3 or TUNEL staining, and found the apoptotic signal remained unchanged after glial ferritin knockdown (Figure 3-figure supplement 3A-D). Furthermore, we overexpressed the baculovirus protein P35 to inhibit apoptosis in glial cells, but failed to restore the lost NBs (Figure 3-figure supplement 3E). Given that glial morphology is important for NB lineages (Speder & Brand, 2018), we used GFP driven by repo-GAL4 to denote glia cells and found that glial chamber was almost intact after glial ferritin knockdown (Figure 3-figure supplement 3F). Together, we conclude that it is iron deficiency in NBs, rather than iron overload in glial cells, that leads to NB loss induced by ferritin defects.

Glial ferritin defects result in impaired Fe-S cluster activity and ATP production

To further prove that defective iron transport leads to the iron deficiency in NBs, we examined the iron availability by measuring aconitase activity in the brain. The activity of aconitase is sensitive to the level of Fe-S synthesis, so it can serve as an indicator for the availability of iron in the cell (Haile et al., 1992; Tong & Rouault, 2006). As shown in the Figure 4A, aconitase activity of ferritin knockdown diminished significantly, indicating that the availability of iron in the brain was reduced. This result also implies Fe-S clusters in NBs are probably deficient when knocking down ferritin in glia. Fe-S clusters are essential to activate the aconitase that converts citrate into isocitrate in the tricarboxylic acid cycle (TCA cycle) in mitochondria, and are one type of critical electron carrier for electron transport chain (ETC), especially for the Complex I, II and III function. Considering that TCA cycle and ETC are critical biological processes for ATP production, we proposed that knockdown of glial ferritin might lead to ATP deficiency due to the block of TCA cycle and ETC, which resulted in the failure of maintaining NBs and their proliferation. To verify this hypothesis, we first assessed the function of Fe-S clusters in NBs by knocking down the components of the cytosolic iron-sulfur protein assembly (CIA) complex. After knocking down the CIA component CG17683, galla2, and ciao1 in NBs respectively, the proliferation level of NBs decreased significantly and brain size became smaller (Figure 4B and C), which phenocopied the knockdown of Fer1HCH or Fer2LCH in glia. This result implies that Fe-S clusters are indispensable for NBs to maintain their proliferation.

Glial ferritin defects result in impaired Fe-S cluster activity and ATP production

(A) The determination of cytosolic aconitase activity.

(B) Knockdown of CIA complex in NBs.

(C) Quantification of NB proliferation in (B).

(D) The schematic diagram of NB sorting procedure.

(E) Venn diagram of downregulated genes after glial ferritin knockdown.

(F) GO enrichment of down-regulated genes in glial ferritin knockdown comparing with control.

(G) KEGG pathway enrichment of down-regulated genes after glial ferritin knockdown.

(H) NAD+ level was indicated by SoNar.

(I) Quantification of normalized SoNar signal in (H).

(J) The determination of ATP level in CNS.

(K) NB number in glial ferritin knockdown was rescued by Ndi1 overexpression in NBs.

(L) Quantification of NB number in (K).

(A and J, n=3; C, n=5,6,6,9; I, n=37,42; L, n=7,7,6,6,7; *: p<0.05; **: p<0.01; ***: p<0.001; ****: p<0.0001)

To identify the biological processes related to Fe-S clusters in NB lineages, we performed transcriptome analysis of sorted NB lineages from glial ferritin knockdown as well as the control groups. We used NB driver Insc-GAL4 to express UAS-DsRed, labeling NB lineages, and then sorted these cells via fluorescence-activated cell sorting (FACS) (Figure 4D) (Harzer, Berger, Conder, Schmauss, & Knoblich, 2013). We found that thousands of genes were significantly altered in the Fer1HCH and Fer2LCH knockdown groups compared to the control group. Specifically, 880 genes were simultaneously downregulated among the significantly changed genes in both the Fer1HCH and Fer2LCH knockdown groups contrasted to the control group (Figure 4E). Given that iron deficiency in NBs probably leads to the defects of energy production and biosynthesis, we performed gene enrichment analysis in these genes based on Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG). GO term analysis showed that mitochondrial organization, translation, transport, oxidative phosphorylation (OxPhos), and ETC were enriched in the Biological Processes (Figure 4F), indicating NB mitochondrial functions including OxPhos and ETC were abnormal after glial ferritin knockdown. Moreover, NADH dehydrogenase and ribosome related terms were also enriched whether in the Cellular Component or Molecular Function (Figure 4F), suggesting that NADH dehydrogenase and biosynthesis were defective. Further, we validated the enriched term associated with mitochondrial function by qRT-PCR (Figure 4-figure supplement 1A). And the enriched KEGG pathways included OxPhos and ribosome (Figure 4G), aligning with the GO enrichment results and supporting the hypothesis that glial ferritin knockdown disrupted mitochondrial function including OxPhos and ETC as well as further biosynthesis.

To confirm whether the function of NADH dehydrogenase was disrupted, we used the SoNar that could bind to NAD+ with specific conformation and excitatory wavelength to monitor the NAD+ level (Bonnay et al., 2020). SoNar was driven by the Insc-GAL4 to indicate the NAD+ level in NBs. Considering the protein accumulation in NBs due to the lower proliferation rate in glial ferritin knockdown group, the signal of SoNar was normalized to DsRed Stinger driven by Insc-GAL4. We observed significantly declined SoNar/DsRed after knocking down Fer2LCH (Figure 4H and I), suggesting the normalized level of NAD+ in NBs was decreased, which supports the conclusion that the activity of NADH dehydrogenase was declined. To further verify the change of energy production, we utilized the ATP Determination Kit which offers a convenient bioluminescence assay for quantitative determination of ATP with recombinant firefly luciferase and its substrate D-luciferin. We found that ATP level decreased significantly after glial ferritin knockdown when compared to the control (Figure 4J). Therefore, we tried to supply the energy for NBs by restoring bioenergy-dependent NAD+ metabolism. As expected, the NB number could be partially rescued by overexpressing proton-pump-independent yeast mitochondrial NADH dehydrogenase Ndi1 in NBs under the background of glial ferritin knockdown (Figure 4K and L) (Bonnay et al., 2020). In summary, the iron deficiency leads to NB loss due to the insufficient energy required for essential biological processes.

Glial ferritin maintains NBs via preventing premature differentiation

To investigate how glial ferritin knockdown leads to NB loss, we firstly examined whether NB origin was affected via altering the timing of ferritin knockdown. We used the temperature-sensitive GAL4 inhibitor GAL80ts to suppress ferritin RNAi at a permissive temperature (18℃) during embryonic stage, and then shifted to a restrictive temperature (29℃), or vice versa until we dissected them at the third-instar larval stage. Ferritin knockdown during the larval stage led to severe brain development defects (Figure 5A and B), while the knockdown only during the embryonic stage showed no apparent abnormality (Figure 5-figure supplement 1A and B). Together, these results indicate that glial ferritin is required for NB maintenance at the larval stage. Consistent with this result, we found no obvious defects of brain at the first-instar larval stage (0-4 hours after larval hatching) when knocking down glial ferritin (Figure 5-figure supplement 1C). Since newborn NBs are delaminated from the neuroectoderm during the embryonic stage (stage 9 to 11), and ferritin RNAi only during the embryonic stage did not show any phenotype, suggesting that NB loss after glial ferritin knockdown is not ascribed to defective NB origin.

Glial ferritin defects lead to the premature differentiation of NBs

(A) Temporal window of glial ferritin maintaining NBs. The temperature shift of flies with glial ferritin knockdown from the permissive temperature (18℃) to the restrictive temperature (29℃).

(B) Quantification of NB number and proliferation in (A).

(C) Pros staining of knocking down glial ferritin.

(D) Quantification of ratio of Pros+Dpn+ NBs to Dpn+ NBs.

(E) Knockdown of pros in NBs rescues the NB number.

(F) The ratio of brain with NB loss in (E).

(B, n=6, 5, 6; n=6, 5, 4; D, n=5; F, n=3; *: p<0.05; ***: p<0.001; ****: p<0.0001)

As mentioned before, we did not detect any apoptotic signal in the larval brain after glial ferritin knockdown, suggesting that there is no association between apoptosis and NB loss. The double binary systems, lexA/lexAop and GAL4/UAS, were employed to further exclude the possibility of apoptosis. We overexpressed P35 in NBs by Insc-GAL4 and knocked down ferritin by repo-lexA, which failed to restore the decreased number of NBs (Figure 5-figure supplement 1D). These results reinforce the notion that NB loss occurs independent of the apoptotic pathway.

Next, we hypothesized that the NB loss might be attributed to nuclear-Prospero (Pros)-dependent premature differentiation of NBs (Cabernard & Doe, 2009; Choksi et al., 2006). Pros in NB nuclear was hardly detected in third-instar WT larvae, but was increased significantly with glial ferritin knockdown (Figure 5C and D). To further verify that the NB loss caused by glial ferritin knockdown was due to nuclear-Pros-dependent premature differentiation, we used the Insc-GAL4 to drive UAS-pros-RNAi in order to prevent the NB loss in glial ferritin knockdown background. We found that the NB number was substantially restored (Figure 5E and F). Interestingly, despite of dramatic NB number increase, the decreased proliferation rate was not rescued. We proposed that pros knockdown in NBs blocked premature terminal differentiation, but it could not replenish the reduced energy caused by glial ferritin knockdown. As NB self-renewal and proliferation are both energy-consuming processes, the self-renewal was rescued via inhibiting differentiation by Insc-GAL4 driving UAS-pros-RNAi, but at the expense of energy consumption for proliferation, leaving the proliferation rate unchanged, or even lower. Together, we conclude that glial ferritin maintains the number of NBs by preventing the nuclear-Pros-dependent premature differentiation.

Given that reduced proliferation is a distinctive characteristic of quiescent NBs, we tried to determine whether NBs were in a quiescent state after glial ferritin knockdown by examining the other two key features associated with quiescent NBs: decreased cell size and the formation of the extended cellular protrusions (Chell & Brand, 2010; Ly et al., 2019). However, we found that the NB size remained at the same level as control, and the extended cellular protrusion was not observed after glial ferritin knockdown (Figure 2B and Figure 5-figure supplement 2A). Furthermore, since insulin signaling is necessary and sufficient to reactivate NBs from quiescence (Sousa-Nunes, Yee, & Gould, 2011), we tried to overexpress a constitutively active form of InR (InRact) in NBs to promote reactivation at the background of glial ferritin knockdown. However, InRact overexpression failed to ameliorate the NB abnormalities including the decreased proliferation (Figure 5-figure supplement 2B and C). In conclusion, NBs exhibited a low proliferation rate after glial ferritin knockdown, but were not in a quiescent state.

Ferritin functions as a potential target for tumor suppression

We have demonstrated that glial ferritin is essential for NB self-renewal and proliferation by providing glia-to-NB iron transport to support energy production in NBs. We next tried to determine whether manipulating this biological function could suppress tumor development. We induced tumor development by driving UAS-brat-RNAi or UAS-numb-RNAi under the control of Insc-GAL4, and knocked down ferritin genes under the control of repo-lexA. We observed that the size of tumors was strikingly decreased in the larval brains lacking glial ferritin (Figure 6A and B). This result suggests that glial ferritin is required for tumor development, highlighting a potential target for tumor treatment.

Ferritin functions as a potential target for tumor suppression

(A) Knockdown of ferritin in glia could inhibit the tumor induced by brat RNAi or numb RNAi.

(B) Quantification of tumor size in (A).

(C) Iron chelator BPS supplemented in the food suppressed brain tumor growth.

(D) Quantification of tumor size in (C).

(E) Bioluminescence signal in mice with glioma on day 14.

(F) Quantification of bioluminescence signal in (E).

(G) Kaplan-Meier survival curve of mice treated with DFP.

(B, n=10,7,7,5; D, n=8,6,6,6; F, n=6,7; G, n=7,6; *: p<0.05; ***: p<0.001; ****: p<0.0001)

To verify whether iron is involved in the tumor suppression induced by blocking glial ferritin, we tried to decrease iron level by adding the iron chelator BPS in the Drosophila food. Tumor formation in the brain at the third instar larval stage was induced with the same method as mentioned above. The addition of BPS into dietary food significantly reduced the tumor size (Figure 6C and D), suggesting that iron is required for tumor development in Drosophila larval brain. Considering the conservation of iron’s fundamental function as an essential nutrient for cell growth and proliferation, we thus tried to suppress brain tumor in mice via iron deficiency. Initially, we constructed orthotopic glioma model in C57BL/6J mice. One week after inoculation of glioma cells (GL261-luc), the iron chelator DFP was intraperitoneally injected every two days based on the murine weight until being dissected on day 7, 14, and 21, respectively. Histological examination with HE staining revealed that the size of glioma in mice, injected with DFP, decreased significantly on day 14 and 21 (Figure 6-figure supplement 1A-C). Since glioma cells GL261-luc carry the luciferase gene, we injected the substrate luciferin potassium salt into the tumor-bearing mice and detected the bioluminescence signal of mouse gliomas using IVIS Spectrum. This result showed that the bioluminescence signal of glioma in mice with the injection of 10 mg/ml DFP was significantly reduced (Figure 6E and F). The above results suggested that iron deficiency induced by DFP could suppress glioma in mice effectively. Furthermore, we assessed the survival time of the tumor-bearing mice with DFP or saline injection and found that 10 mg/ml DFP injection could significantly prolong the survival time of tumor-bearing mice (Figure 6G). Altogether, blocking iron into tumor is a potential therapy for tumor suppression.

Glial ferritin level is regulated by NBs

The aforementioned results have highlighted the significant role of glial ferritin in controlling NB number and proliferation. We next wonder whether NBs could provide feedback to regulate glial ferritin level. To answer this question, we analyzed ferritin level after altering NB proliferation. Previous studies have corroborated that NB proliferation was diminished by blocking Tor signaling pathway or enhanced by knocking down tumor suppressor genes, such as brat or numb (Bowman et al., 2008; Reichardt et al., 2018; Sousa-Nunes et al., 2011). When NBs were in low proliferative status, Fer2LCH subunits in the brain were significantly decreased, while Fer1HCH subunits declined slightly with no significance (Figure 7A-D). Conversely, highly proliferative NBs, able to form tumors, induced significantly increased level of the ferritin (Figure 7A-D). In summary, the production of ferritin in glial cells depends on the level of NB proliferation.

The level of ferritin is controlled by NB proliferation

(A) Fer1HCH subunit level in Drosophila CNS after manipulating proliferation.

(B) Fer2LCH subunit level in Drosophila CNS after manipulating proliferation.

(C) Quantification of Fer1HCH level in (A).

(D) Quantification of Fer2LCH level in (B).

(E) The model of glial ferritin regulating NBs.

(C and D, n=3; ns, not significant; *: p<0.05; **: p<0.01; ***: p<0.001)

Discussion

Here we demonstrate that glia regulates iron metabolism in the brain to sustain NBs and their proliferation. Glial ferritin coordinates Zip13 to provide iron for NBs. Once ferritin in glia is defective, NBs proliferate much slower and differentiate prematurely, implying that iron is a necessary metallic element for NBs. Iron participates in the synthesis of Fe-S clusters which activate aconitase required for TCA and facilitate electron-transfer processes involved in OxPhos. Since TCA and OxPhos are essential steps of ATP production, glial ferritin supplies iron into NBs to ensure energy production for maintaining NBs and their proliferation (Figure 7E).

NSCs and glial niche

Previous literatures have reported how glial niche regulates NB reactivation, proliferation, and survival. Blood-brain barrier (BBB) glia secretes dILP2 and dILP6 that bind on InR of NBs, which switches on Insulin signaling to reactivate NBs (Chell & Brand, 2010; Speder & Brand, 2014). In addition, Pvr (PDGFR ortholog) signaling in cortex glia sustains NB survival and proliferation in a PI3K-DE-cadherin-dependent manner (Read, 2018). However, it is unknown whether glial niche is required for maintaining the self-renewal of NBs. In this study, we demonstrate that glial niche maintains the self-renewal and proliferation of NBs through supplying the necessary microelement iron into NBs for enough energy production, which implies that mammalian glia likewise provides critical microelement for maintaining NSC function. Oligodendrocytes have been reported to secret ferritin to protect neuron (Mukherjee et al., 2020), however, it remains unknown whether the ferritin secretion is employed by oligodendrocytes to regulate NSCs. Astrocytes regulate NSC proliferation via secreting growth factors (Shetty, Hattiangady, & Shetty, 2005), but it is unclear whether ferritin from astrocytes regulates NSC proliferation and other NSC behaviors, such as self-renewal and differentiation. Based on our findings, it is feasible to regulate the self-renewal and differentiation of NSCs indirectly through niche cells, which would help to reduce the risk of directly manipulating NSCs to treat neural diseases.

Ferritin and iron

Our results show that ferritin generated outside the brain was not detected in the CNS and vice versa (data not shown), suggesting ferritin can hardly cross the BBB glia, possibly because ferritin is too large to permeate. In this way, BBB glia separates the labile systemic iron from local iron inside the brain to guarantee the normal brain development, as the neural cells, including NBs, require a stable iron level for continuous energy generation during neurogenesis.

Although the dominant function of ferritin is iron storage in mammal, increasing evidence suggests that ferritin also can transport iron. In mice, ferritin heavy chain is secreted by oligodendrocytes via extracellular vesicles (Mukherjee et al., 2020), which possibly endows ferritin with the function of iron transport. Consistently, previous results showed that ferritin can deliver iron across the cultured BBB (Chiou et al., 2019; Fisher et al., 2007). Our results emphasize the critical role of ferritin in iron transport in the CNS. Future studies would be interesting to investigate whether ferritin is able to transport iron in the mammalian brain.

Ferritin specifically targets on the tumor cells, which has been applied to tumor visualization and targeted drug delivery (Fan et al., 2012; Fan et al., 2018). Our study shows that the block of ferritin or iron can inhibit tumor effectively, which offers an explanation that ferritin targets on tumor is to transport iron for maintaining the cell proliferation, highlighting a potential target for tumor suppression.

Iron is an essential component for Fe-S clusters which participate in TCA and OxPhos, providing energy required for self-renewal and proliferation of NBs. Previous study claimed that OxPhos is required for NB proliferation (van den Ameele & Brand, 2019). However, another research observed that OxPhos is dispensable in type II NBs (Bonnay et al., 2020). Our data imply that energy deficiency in NBs caused by iron shortage leads to NB loss and low proliferation possibly through simultaneous disruption of TCA and OxPhos. Consistently, iron-loaded ferritin nanoparticles can be used to improve the self-renewal ability and differentiation potential of human NSCs in vitro (Lee, Yang, Cho, & Cho, 2018). Additionally, previous study has shown that iron-loaded ferritin can promote the proliferation of cultured cell from Drosophila (Li, 2010). Furthermore, iron depletion was reported to suppress human cancer cell growth at G1/S by inducing cyclin D1 proteolysis (Nurtjahja-Tjendraputra, Fu, Phang, & Richardson, 2007). These studies suggest that iron, as a necessary microelement for energy production, is essential to NSC maintenance and proliferation.

Bicellular iron homeostasis

Iron homeostasis can be regulated systemically by the liver-derived hormone hepcidin and the iron exporter ferroportin, and it is also regulated cellularly by the IRP/IRE system (Hentze, Muckenthaler, Galy, & Camaschella, 2010). Here we propose a bicellular iron homeostasis between NB and glial niche cell, which is distinct from classic cellular and systemic iron homeostasis. The bicellular iron homeostasis requires an elaborate mechanism to guarantee normal neurogenesis.

In a proliferative state, NBs receive iron from glial cells via ferritin to fulfill the energy requirements for proliferation and self-renewal. Therefore, the iron required by NBs is mainly stored in the ferritin of glial cells and allocated as needed. The bicellular iron homeostasis mechanism makes glial cells the main pool for ferritin production, relieving NBs from the need to produce ferritin and transferring the potential risk of iron overload-induced oxidative stress to glial cells. Furthermore, glial cells can subsequently clear oxidative stress through lipid droplets (Bailey et al., 2015), ensuring normal NB function during neural development. It is intriguing to explore whether bicellular iron homeostasis exists in other stem cells and their niches.

When normal NBs are converted into immortal tumor NBs, becoming highly proliferative, they need more nutrients including iron (Bonnay et al., 2020; van den Ameele & Brand, 2019). In response, more ferritin is generated to transport iron from the glial cells to NBs. However, it is unclear how NBs communicate with glial cells to determine the level of ferritin production. It’s interesting to identify the specific cues from NBs that regulate the level of ferritin in future studies. In addition, iron distribution in Drosophila brain is not observed due to low iron concentration and technical limitations in iron detection, which results in a lack of direct evidence for iron trafficking via ferritin.

eLife Assessment

This study, which seeks to identify factors from the glial niche that support and maintain neural stem cells, unveils a novel role for ferritin in this process. Furthermore, the work shows that defects in larval brain development resulting from ferritin knockdown can be attributed to impaired Fe-S cluster activity and ATP production. These findings will be valuable to both oncologists and neurobiologists, though the supporting evidence is currently incomplete.

Reviewer #1 (Public Review):

Summary:

This study unveils a novel role for ferritin in Drosophila larval brain development. Furthermore, it pinpoints that the observed defects in larval brain development resulting from ferritin knockdown are attributed to impaired Fe-S cluster activity and ATP production. In addition, knocking down ferritin genes suppressed the formation of brain tumors induced by brat or numb RNAi in Drosophila larval brains. Similarly, iron deficiency suppressed glioma in the mice model. Overall, this is a well-conducted and novel study.

Strengths:

Thorough analyses with the elucidation of molecular mechanisms.

Weaknesses:

Some of the conclusions are not well supported by the results presented.

Reviewer #2 (Public Review):

Summary:

Zhixin and collaborators have investigated if the molecular pathways present in glia play a role in the proliferation, maintenance, and differentiation of Neural Stem Cells. In this case, Drosophila Neuroblasts are used as models. The authors find that neuronal iron metabolism modulated by glial ferritin is an essential element for Neuroblast proliferation and differentiation. They show that loss of glial ferritin is sufficient to impact on the number of neuroblasts. Remarkably, the authors have identified that ferritin produced in the glia is secreted to be used as an iron source by the neurons. Therefore iron defects in glia have serious consequences in neuroblasts and likely vice versa. Interestingly, preventing iron absorption in the intestine is sufficient to reduce NB number. Furthermore, they have identified Zip13 as another regulator of the process. The evidence presented strongly indicates that loss of neuroblasts is due to premature differentiation rather than cell death.

Strengths:

- Comprenhensive analysis of the impact of glial iron metabolism in neuroblast behaviour by genetic and drug-based approaches as well as using a second model (mouse) for some validations.

- Using cutting-edge methods such as RNAseq as well as very elegant and clean approaches such as RNAi-resistant lines or temperature-sensitive tools

- Goes beyond the state of the art highlighting iron as a key element in neuroblast formation as well as as a target in tumor treatments.

Weaknesses:

Although the manuscripts have clear strengths, there are also some strong weaknesses that need to be addressed.

- Some literature is missing

- In general, the authors succeeded but in some cases, the authors´ claims are not fully supported by the evidence presented and additional experiments are critical to discriminate among different hypotheses.

- Moreover, some potential flaws might be present in the analysis of cell death and mitochondrial iron.

Reviewer #3 (Public Review):

In this manuscript, Ma et al seek to identify stem cell niche factors. They perform an RNAi screen in glial cells and screen for candidates that support and maintain neuroblasts (NBs) in the developing fly brain. Through this, they identify two subunits of ferritin, which is a conserved protein that can store iron in cells in a non-toxic form and release it in a controlled manner when and where required. They present data to support the conclusion that ferritin produced in glia is released and taken up by NBs where it is utilised by enzymes in the Krebs cycle as well as in the electron transport chain. In its absence from glia, NBs are unable to generate sufficient energy for division and therefore prematurely differentiate via nuclear prospero resulting in small brains. The work will be of interest to those interested in neural stem cells and their non-cell autonomous control by niches.

The past decade has seen a growing appreciation of how glial cells support and maintain NBs during development. The authors' discovery of glial-derived ferritin providing essential iron atoms for energy production is interesting and important. They have employed a variety of genetic tools and assays to uncover how ferritin in glia might support NBs. This is particularly challenging because there are no direct ways of assaying for iron or energy consumption in a cell-specific manner.

There are however instances where conclusions are drawn to support the story being developed without considering the equally plausible alternative explanations that should ideally be addressed.

For example, the data supporting the transfer of ferritin from glia to NBs was weak given the misexpression system used; the Shi[ts] experiment was also not convincing (perhaps they have more representative images?).

The iron manipulation experiments are in the whole animal and it is likely that this affects general feeding behaviour, which is known to affect NB exit from quiescence and proliferative capacity. The loss of ferritin in the gut and iron chelators enhancing the NB phenotype are used as evidence that glia provide iron to NB to support their number and proliferation. Since the loss of NB is a phenotype that could result from many possible underlying causes (including low nutrition), this specific conclusion is one of many possibilities.

Similarly, knockdown of the FeS protein assembly components phenocopy glial ferritin knock down. Since iron is so important for the TCA and the ETC, this is not surprising, but the similarities in the two phenotypes seem insufficient to say that it's glial ferritin that's causing the lack of iron in the NB and therefore resulting in loss of NBs.

Pros RNAi will certainly result in an increase in NB numbers because the loss of pros results in an inability of NB progeny to differentiate. This (despite the slight increase in nuclear pros) is not sufficient to infer that glial ferritin knockdown results in premature differentiation of NBs via nuclear pros.

I recognise these are challenging to prove irrefutably, however, the frequency of such expansive interpretations of data is of concern.

Author response:

The following is the authors’ response to the original reviews.

Reviewer #3 (Public Review):

The iron manipulation experiments are in the whole animal and it is likely that this affects general feeding behaviour, which is known to affect NB exit from quiescence and proliferative capacity. The loss of ferritin in the gut and iron chelators enhancing the NB phenotype are used as evidence that glia provide iron to NB to support their number and proliferation. Since the loss of NB is a phenotype that could result from many possible underlying causes (including low nutrition), this specific conclusion is one of many possibilities.

We have investigated the feeding behavior of fly by Brilliant Blue (sigma, 861146)[1]. Our result showed that the amount of dye in the fly body were similar between control group and BPS group, suggesting that BPS almost did not affect the feeding behavior (Figure 3—figure supplement 1A).

Recommendations for the authors:

Reviewer #1 (Recommendations For The Authors):

There was a gap between the Pros nuclear localization and downstream targets of ferritin, particularly NADH dehydrogenase and biosynthesis. Could overexpression of Ndi1 restore Pros localization in NBs?

Ferritin defect downregulates iron level, which leads to cell cycle arrest of NBs via ATP shortage. And cell cycle arrest of NBs probably results in NB differentiation[2, 3]. We have added the experiment in Figure 5—figure supplement 2. This result showed that overexpression of Ndi1 could significantly restore Pros localization in NBs.

The abstract requires revision to cover the major findings of the manuscript, particularly the second half.

We revised the abstract to add more major findings of the manuscript in the second half as follows:

“Abstract

Stem cell niche is critical for regulating the behavior of stem cells. Drosophila neural stem cells (Neuroblasts, NBs) are encased by glial niche cells closely, but it still remains unclear whether glial niche cells can regulate the self-renewal and differentiation of NBs. Here we show that ferritin produced by glia, cooperates with Zip13 to transport iron into NBs for the energy production, which is essential to the self-renewal and proliferation of NBs. The knockdown of glial ferritin encoding genes causes energy shortage in NBs via downregulating aconitase activity and NAD+ level, which leads to the low proliferation and premature differentiation of NBs mediated by Prospero entering nuclei. More importantly, ferritin is a potential target for tumor suppression. In addition, the level of glial ferritin production is affected by the status of NBs, establishing a bicellular iron homeostasis. In this study, we demonstrate that glial cells are indispensable to maintain the self-renewal of NBs, unveiling a novel role of the NB glial niche during brain development.”

In Figure 2B Mira appeared to be nuclear in NBs, which is inconsistent with its normal localization. Was it Dpn by mistake?

In Figure 2B, we confirmed that it is Mira. Moreover, we also provide a magnified picture in Figure 2B’, showing that the Mira mainly localizes to the cortex or in the cytoplasm as previously reported.

Figure 2C, Fer1HCH-GFP/mCherry localization was non-uniform in the NBs revealing 1-2 regions devoid of protein localization potentially corresponding to the nucleus and Mira crescent enrichment. It is important to co-label the nucleus in these cells and discuss the intracellular localization pattern of Ferritin.

We have revised the picture with nuclear marker DAPI in Figure 2C. The result showed that Fer1HCH-GFP/Fer2LCH-mCherry was not co-localized with DAPI, which indicated that Drosophila ferritin predominantly distributes in the cytosol[4, 5]. As for the concern mentioned by this reviewer, GFP/mCherry signal in NBs was from glial overexpressed ferritin, which probably resulted in non-uniform signal.

In Figure 3-figure supplement 3F, glial cells in Fer1HCH RNAi appeared to be smaller in size. This should be quantified. Given the significance of ferritin in cortex glial cells, examining the morphology of cortex glial cells is essential.

In Figure 3—figure supplement 3F, we did not label single glial cells so it was difficult to determine whether the size was changed. However, it seems that the chamber formed by the cellular processes of glial cells becomes smaller in Fer1HCH RNAi. The glial chamber will undergo remodeling during neurogenesis, which responses to NB signal to enclose the NB and its progeny[6]. Thus, the size of glial chamber is regulated by NB lineage size. In our study, ferritin defect leads to the low proliferation, inducing the smaller lineage of each NB, which likely makes the chamber smaller.

Since the authors showed that the reduced NB number was not due to apoptosis, a time-course experiment for glial ferritin KD is recommended to identify the earliest stage when the phenotype in NB number /proliferation manifests during larval brain development.

We observed brains at different larval stages upon glial ferritin KD. The result showed that NB proliferation decreased significantly, but NB number declined slightly at the second-instar larval stage (Figure 5—figure supplement 1E and F), suggesting that brain defect of glial ferritin KD manifests at the second-instar larval stage.

Transcriptome analysis on ferritin glial KD identified genes in mitochondrial functions, while the in vivo EM data suggested no defects in mitochondria morphology. A short discussion on the inconsistency is required.

For the observation of mitochondria morphology via the in vivo EM data, we focused on visible cristae in mitochondria, which was used to determine whether the ferroptosis happens[7]. It is possible that other details of mitochondria morphology were changed, but we did not focus on that. To describe this result more accurately, we replaced “However, our observation revealed no discernible defects in the mitochondria of NBs after glial ferritin knockdown” with the “However, our result showed that the mitochondrial double membrane and cristae were clearly visible whether in the control group or glial ferritin knockdown group, which suggested that ferroptosis was not the main cause of NB loss upon glial ferritin knockdown” in line 207-209.

The statement “we found no obvious defects of brain at the first-instar larval stage (0-4 hours after larval hatching) when knocking down glial ferritin (Figure 5-figure supplement 1C).” lacks quantification of NB number and proliferation, making it challenging to conclude.

We have provided the quantification of NB number and proliferation rate of the first-instar larval stage in Figure 5—figure supplement 1C and D. The data showed that there is no significant change in NB number and proliferation rate when knocking down ferritin, suggesting that no brain defect manifests at the first-instar larval stage.

A wild-type control is necessary for Figure 6A-C as a reference for normal brain sizes.

We have added Insc>mCherry RNAi as a reference in Figure 6A-D, which showed that the brain size of tumor model is larger than normal brain. Moreover, we removed brat RNAi data from Figure 6A-D to Figure 6—figure supplement 1A-D for the better layout.

In Figures 6B, D, “Tumor size” should be corrected to “Larval brain volume”.

Here, we measured the brain area to assess the severity of the tumor via ImageJ instead of 3D data of the brain volume. So we think it would be more appropriate to use the “Larval brain size” than “Larval brain volume” here. Thus, we have corrected “Tumor size” to “Larval brain size” in Figure 6B and D to Figure 6—figure supplement 1B and D.

Considering that asymmetric division defects in NBs may lead to premature differentiation, it is advisable to explore the potential involvement of ferritin in asymmetric division.

aPKC is a classic marker to determine the asymmetric division defect of NB. We performed the aPKC staining and found it displayed a crescent at the apical cortex based on the daughter cell position whether in control or glial ferritin knockdown (Figure 5—figure supplement 3A). This result indicated that there was no obvious asymmetric defect after glial ferritin knockdown.

In the statement "Secondly, we examined the apoptosis in glial cells via Caspase-3 or TUNEL staining, and found the apoptotic signal remained unchanged after glial ferritin knockdown (Figure 3-figure supplement 3A-D).", replace "the apoptosis in glial cells" with "the apoptosis in larval brain cells".

We have replaced "the apoptosis in glial cells" with "the apoptosis in larval brain cells" in line 216.

Include a discussion on the involvement of ferritin in mammalian brain development and address the limitations associated with considering ferritin as a potential target for tumor suppression.

We have added the discussion about ferritin in mammalian brain development in line 428-430 and limitation of ferritin for suppressing tumor in line 441-444.

Indicate Insc-GAL4 as BDSC#8751, even if obtained from another source. Additionally, provide information on the extensively used DeRed fly stock used in this study within the methods section.

We provided the stock information of Insc-GAL4 and DsRed in line 673-674.

Reviewer #2 (Recommendations For The Authors):

Major points:

The number of NBs differs a lot between experiments. For example, in Fig 1B and 1K controls present less than 100 NBs whereas in Figure 1 Supplementary 2B it can be seen that controls have more than 150. Then, depending on which control you compare the number of NBs in flies silencing Fer1HCH or Fer2LCH, the results might change. The authors should explain this.

Figure 1 Supplementary 2B (Figure 1 Supplementary 3B in the revised version) shows NB number in VNC region while Fig 1B and 1K show NB number in CB region. At first, we described the general phenotype showing the NB number in CB and VNC respectively (Fig 1 and Fig 1-Supplementary 1 and 3 in the revised version). And the NB number is consistent in each region. After then, we focused on NB number in CB for the convenience.

This reviewer encourages the authors to use better Gal4 lines to describe the expression patterns of ferritins and Zip13 in the developing brain. On the one hand, the authors do not state which lines they are using (including supplementary table). On the other hand, new Trojan GAL4 (or at least InSite GAL4) lines are a much better tool than classic enhancer trap lines. The authors should perform this experiment.

All stock source and number were documented in Table 2. Ferritin GAL4 and Zip13 GAL4 in this study are InSite GAL4. In addition, we also used another Fer2LCH enhancer trapped GAL4 to verify our result (DGRC104255) and provided the result in Figure 2—figure supplement 1. Our data showed that DsRed driven by Fer2LCH-GAL4 was co-localized with the glia nuclear protein Repo, instead of the NB nuclear protein Dpn, which was consistent with the result of Fer1HCH/Fer2LCH GAL4. In addition, we will try to obtain the Trojan GAL4 (Fer1HCH/Fer2LCH GAL4 and Zip13 GAL4) and validate this result in the future.

The authors exclude very rapidly the possibility of ferroptosis based only on some mitochondrial morphological features without analysing the other hallmarks of this iron-driven cell death. The authors should at least measure Lipid Peroxidation levels in their experimental scenario either by a kit to quantify by-products of lipid peroxidation such as Malonaldehide (MDA) or using an anti 4-HNE antibody.

We combined multiple experiments to exclude the possibility of ferroptosis. Firstly, ferroptosis can be terminated by iron chelator. And we fed fly with iron chelator upon glial ferritin knockdown, but NB number and proliferation were not restored, which suggested that ferroptosis probably was not the cause of NB loss induced by glial ferritin knockdown (Figure 3B and C). Secondly, Zip13 transports iron into the secretary pathway and further out of the cells in Drosophila gut[8]. Our data showed that knocking down iron transporter Zip13 in glia resulted in the decline of NB number and proliferation, which was consistent with the phenotype upon glial ferritin knockdown (Figure 3E-G). More importantly, the knockdown of Zip13 and ferritin simultaneously aggravated the phenotype in NB number and proliferation (Figure 3E-G). These results suggested that the phenotype was induced by iron deficiency in NB, which excluded the possibility of iron overload or ferroptosis to be the main cause of NB loss upon glial ferritin knockdown. Finally, we observed mitochondrial morphology on double membrane and the cristae that are critical hallmarks of ferroptosis, but found no significant damage (Figure 3-figure supplement 2E and F).

In addition, we have added the 4-HNE determination in Figure 3—figure supplement 2G and H. This result showed that 4-HNE level did not change significantly, suggesting that lipid peroxidation was stable, which supported to exclude the possibility that the ferroptosis led to the NB loss upon glial ferritin knockdown.

All of the above results together indicate that ferroptosis is not the cause of NB loss after ferritin knockdown.

A major flaw of the manuscript is related to the chapter Glial ferritin defects result in impaired Fe-S cluster activity and ATP production and the results displayed in Figure 4. The authors talk about the importance of FeS clusters for energy production in the mitochondria. Surprisingly, the authors do not analyse the genes involved in this process such as but they present the interaction with the cytosolic FeS machinery that has a role in some extramitochondrial proteins but no role in the synthesis of FeS clusters incorporated in the enzymes of the TCA cycle and the respiratory chain. The authors should repeat the experiments incorporating the genes NSF1 (CG12264), ISCU(CG9836), ISD11 (CG3717), and fh (CG8971) or remove (or at least rewrite) this entire section.

Thanks for this constructive advice and we have revised this in Figure 4B and C. We repeated the experiment with blocking mitochondrial Fe-S cluster biosynthesis by knocking down Nfs1 (CG12264), ISCU(CG9836), ISD11 (CG3717), and fh (CG8971), respectively. Nfs1 knockdown in NB led to a low proliferation, which was consistent with CIA knockdown. However, we did not observe the obvious brain defect in ISCU(CG9836), ISD11 (CG3717), and fh (CG8971) knockdown in NB. Our interpretation of these results is that Nfs1 probably is a necessary core component in Fe-S cluster assembly while others are dispensable[9].

The presence and aim of the mouse model Is unclear to this reviewer. On the one hand, It Is not used to corroborate the fly findings regarding iron needs from neuroblasts. On the other hand, and without further explanation, authors migrate from a fly tumor model based on modifying all neuroblasts to a mammalian model based exclusively on a glioma. The authors should clarify those issues.

Although iron transporter probably is different in Drosophila and mammal, iron function is conserved as an essential nutrient for cell growth and proliferation from Drosophila to mammal. The data of fly suggested that iron is critical for brain tumor growth and thus we verified this in mammalian model. Glioma is the most common form of central nervous system neoplasm that originates from neuroglial stem or progenitor cells[10]. Therefore, we validated the effect of iron chelator DFP on glioma in mice and found that DFP could suppress the glioma growth and further prolong the survival of tumor-bearing mice.

Minor points

Although referred to adult flies, the authors did not include either in the introduction or in the discussion existing literature about expression of ferritins in glia or alterations of iron metabolism in fly glia cells (PMID: 21440626 and 25841783, respectively) or usage of the iron chelator DFP in drosophila (PMID: 23542074). The author should check these manuscripts and consider the possibility of incorporating them into their manuscript.

Thanks for your remind. We have incorporated all recommended papers into our manuscript line 65-67 and 168.

The number of experiments in each figure is missing.

All experiments were repeated at least three times. And we revised this in Quantifications and Statistical Analysis of Materials and methods.

If graphs are expressed as mean +/- sem, it is difficult to understand the significance stated by the authors in Figure 2E.

We apologize for this mistake and have revised this in Quantifications and Statistical Analysis. All statistical results were presented as means ± SD.

When authors measure aconitase activity, are they measuring all (cytosolic and mitochondrial) or only one of them? This is important to better understand the experiments done by the authors to describe any mitochondrial contribution (see above in major points).

In this experiment, we were measuring the total aconitase activity. We also tried to determine mitochondrial aconitase but it failed, which was possibly ascribed to low biomass of tissue sample.

In this line, why do controls in aconitase and atp lack an error bar? Are the statistical tests applied the correct ones? It is not the same to have paired or unpaired observations.

It is the normalization. We repeated these experiments at least three times in different weeks respectively, because the whole process was time-consuming and energy-consuming including the collection of brains, protein determination and ATP or aconitase determination. And the efficiency of aconitase or ATP kit changed with time. We cannot control the experiment condition identically in different batches. Therefore, we performed normalization every time to present the more accurate result. The control group was normalized as 1 via dividing into itself and other groups were divided by the control. This normalized process was repeated three times. Therefore, there is no error bar in the control group. We think it is appropriate to apply ANOVA with a Bonferroni test in the three groups.

In some cases, further rescue experiments would be appreciated. For example, expression of Ndi restores control NAD+ levels or number of NBs, it would be interesting to know if this is accompanied by restoring mitochondrial integrity and its ability to produce ATP.

We have determined ATP production after overexpressing Ndi1 and provided this result in Figure 4—figure supplement 1B. The data showed that expression of Ndi1 could restore ATP production upon glial Fer2LCH knockdown, which was consistent with our conclusion.

Lines 293-299 on page 7 are difficult to understand.

According to our above results, the decrease of NB number and proliferation upon glial ferritin knockdown (KD) was caused by energy deficiency. As shown in the schematic diagram (Author response image 1), “T” represented the total energy which was used for NB maintenance and proliferation. “N” indicated the energy for maintaining NB number. “P” indicated the energy for NB proliferation. “T” is equal to “N” plus “P”. When ferritin was knocked down in glia, “T”, “N” and “P” declined in “Ferritin KD” compared to “wildtype (WT)”. Knockdown of pros can prevent the differentiation of NB, but it cannot supply the energy for NB, which probably results in the rescue of NB number but not proliferation. Specifically, NB number increased significantly in “Ferritin KD Pros KD” compared to “Ferritin KD”, which resulted in consuming more energy for NB maintenance in “Ferritin KD Pros KD”. As shown in the schematic diagram, “T” was not changed between “Ferritin KD Pros KD” and “Ferritin KD”, whereas ”N” was increased in “Ferritin KD Pros KD” compared to “Ferritin KD”. Thus, “P” was decreased, which suggested that less energy was remained for proliferation, leading to the failure of rescue in NB proliferation. It seemed that the level of proliferation in “Ferritin KD Pros KD” was even lower than “Ferritin KD”.

The schematic diagram of relationship between energy and NB function in different groups.

“T” represents total energy for NB maintenance and proliferation. “N” represents the energy for NB maintenance. “P” represents the energy for NB proliferation. T=N+P

Line 601 should indicate that Tables 2 and 3 are part of the supplementary material.

We have revised this in line 678.

Figure 4-supplement 1. Only validation of 2 genes from a RNAseq seems too little.

We dissected hundreds of brains for sorting NBs because of low biomass of fly brain. This is a difficult and energy-consuming work. Most NBs were used for RNA-seq, so we can only use a small amount of sample left for validation which is not enough for more genes.

Figure 6E, the authors indicate that 10 mg/ml DFP injection could significantly prolong the survival time. Which increase in % is produced by DFP?

We have provided the bar graph in Author response image 2. The increase is about 16.67% by DFP injection.

The bar graph of survival time of mice treated with DFP.

(The unpaired two-sided Student’s t test was employed to assess statistical significance. Statistical results were presented as means ± SD. n=7,6; *: p<0.05)

Reviewer #3 (Recommendations For The Authors):

As I read the initial results that built the story (glia make ferritin>release it> NBs take them up>use it for TCA and ETC) I kept thinking about what it meant for NBs to be 'lost'. This led me to consider alternate possibilities that the results might point to, other than the ones the authors were suggesting. It was only in Figure 5 that the authors ruled out some of those possibilities. I would suggest that they first illustrate how NBs are lost upon glial ferritin loss of function before they delve into the mechanism. This would also be a place to similarly address that glial numbers and general morphology are unchanged upon ferritin loss.

This recommendation provides a valuable guideline to build this story especially for researchers who are interested in neural stem cell studies. Actually, we tried this logic to present our study but found that there are several gaps in the middle of the manuscript, such as the relationship between glial ferritin and Pros localization in NB, so that the whole story cannot be fluently presented. Therefore, we decided to present this study in the current way.

More details of the screen would be useful to know. How many lines did they screen, what was the assay? This is not mentioned anywhere in the text.

We have added this in Screen of Materials and methods. We screened about 200 lines which are components of classical signaling pathways, highly expressed genes in glial cells or secretory protein encoding genes. UAS-RNAi lines were crossed with repo-Gal4, and then third-instar larvae of F1 were dissected. We got the brains from F1 larvae and performed immunostaining with Dpn and PH3. Finally, we observed the brain in Confocal Microscope.

Many graphs seem to be repeated in the main figures and the supplementary data. This is unnecessary, or at least should be mentioned.

We appreciate your kind reminder. However, we carefully went through all the figures and did not find the repeated graphs, though some of them look similar.

The authors mention that they tested which glial subtypes ferritin is needed in, but don't show the data. Could they please show the data? Same with the other iron transport/storage/regulation. Also, in both this and later sections, the authors could mention which Gal4 was used to label what cell types. The assumption is that the reader will know this information.

We have added the result of ferritin knockdown in glial subpopulations in Figure 1—figure supplement 2. However, considering that the quantity of iron-related genes, we did not take the picture, but we recorded this in Table 3.

For all their images showing colocalisation, magnified, single-colour images shown in grayscale will be useful. For example, without the magnification, it is not possible to see the NB expression of the protein trap line in Figure 2B. A magnified crop of a few NBs (not a single one like in 2C) would be more useful.

We have provided Figure 2A’, B’, D’ and Figure 3D’ as suggested.

There are a lot of very specific assays used to detect ROS, NAD, aconitase activity, among others. It would be nice to have a brief but clear description of how they work in the main text. I found myself having to refer to other sources to understand them. (I believe SoNAR should be attributed to Zhao et al 206 and not Bonnay et al 2020.)

We have added a brief description about ROS, aconitase activity, NAD in line 198-199, 229-231, and 269 as suggested.

I did not understand the normalisation done with respect to SoNAR. Is this standard practice? Is the assumption that 'overall protein levels will be higher in slowly proliferating NBs' reasonable? This is why they state the need to normalise.

The SoNAR normalization is not a standard practice. However, we think that our normalization of SoNar is reasonable. According to our results, the expression level of Dpn and Mira seemed higher in glial ferritin knockdown, so we speculated that some proteins accumulated in slowly proliferating NBs. Thus, we used Insc-GAL4 to drive DsRed for indicating the expression level of Insc and found that DsRed rose after glial ferritin knockdown, suggesting that Insc expression was increased indeed. Therefore, we have to normalize SoNar driven by Insc-GAL4 based on DsRed driven by Insc-Gal4, which eliminates the effect of increased Insc upon glial ferritin knockdown.

FAC is mentioned as a chelator? But the authors seem to use it oppositely. Is there an error?

FAC is a type of iron salt, which is used to supply iron. We have also indicated that in line 156 according to your advice.

The lack of any cell death in the L3 brain surprised me. There should be plenty of hemilineages that die, as do many NBs, particularly in the abdominal segments. Is the stain working? Related to this, P35 is not the best method for rescuing cell death. H99 might be a better way to go.

We were also surprised to see this result and repeated this experiment for several times with both negative and positive controls. Moreover, we also used TUNEL to validate this result, which led to the same result. We will try to use H99 to rescue NB loss in the future, because it needs to be integrated and recombined with our current genetic tools.

It would be nice to see the aconitase activity signal as opposed to just the quantification.

This method can only determine the absorbance for indicating aconitase activity, so our result is just the quantification.

Glia are born after NBs are specified. In fact, they arise from NBs (and glioblasts). So, it's unlikely that the knockdown of ferritin in glia can at all affect initial NB specification.

We completely agree with this statement.

The section on tumor suppression seems out of place. The fly data on which the authors base this as an angle to chase is weak. Dividing cells will be impaired if they have inadequate energy production. As a therapeutic, this will affect every cell in the body. I'm not sure that cancer therapeutics is pursuing such broadly acting lines of therapies anymore.

Our data suggested that iron/ferritin is more critical for high proliferative cells. Tumor cells have a high expression of TfR (Transferrin Receptor)[11], which can bind to Transferrin and ferritin[12]. And ferritin specifically targets on the tumor cells[11]. Thus, we think iron/ferritin is extremely essential for tumor cells. If we can find the appropriate dose of iron/ferritin inhibitor, suppressing tumor growth but maintaining normal cell growth, iron/ferritin might be an effective target of tumor treatment.

The feedback from NB to glial ferritin is also weak data. The increased cell numbers (of unknown identity) could well be contributing to the increase in ferritin. I would omit the last two sections from the MS.

In brat RNAi and numb RNAi, increased cells are NB-like cells, which cannot undergo further differentiation and are not expected to produce ferritin. More importantly, we used Repo (glia marker) as the reference and quantified the ratio of ferritin level to Repo level, which can exclude the possibility that increased glial cells lead to the increase in ferritin.

References

(1) Tanimura T, Isono K, Takamura T, et al. Genetic Dimorphism in the Taste Sensitivity to Trehalose in Drosophila-Melanogaster. J Comp Physiol, 1982,147(4):433-7

(2) Myster DL, Duronio RJ. Cell cycle: To differentiate or not to differentiate? Current Biology, 2000,10(8):R302-R4

(3) Dalton S. Linking the Cell Cycle to Cell Fate Decisions. Trends in Cell Biology, 2015,25(10):592-600

(4) Nichol H, Law JH, Winzerling JJ. Iron metabolism in insects. Annu Rev Entomol, 2002,47:535-59

(5) Pham DQ, Winzerling JJ. Insect ferritins: Typical or atypical? Biochim Biophys Acta, 2010,1800(8):824-33

(6) Speder P, Brand AH. Systemic and local cues drive neural stem cell niche remodelling during neurogenesis in Drosophila. Elife, 2018,7

(7) Mumbauer S, Pascual J, Kolotuev I, et al. Ferritin heavy chain protects the developing wing from reactive oxygen species and ferroptosis. PLoS Genet, 2019,15(9):e1008396

(8) Xiao G, Wan Z, Fan Q, et al. The metal transporter ZIP13 supplies iron into the secretory pathway in Drosophila melanogaster. Elife, 2014,3:e03191

(9) Marelja Z, Leimkühler S, Missirlis F. Iron Sulfur and Molybdenum Cofactor Enzymes Regulate the Life Cycle by Controlling Cell Metabolism. Front Physiol, 2018,9

(10) Morgan LL. The epidemiology of glioma in adults: a "state of the science" review. Neuro-Oncology, 2015,17(4):623-4

(11) Fan K, Cao C, Pan Y, et al. Magnetoferritin nanoparticles for targeting and visualizing tumour tissues. Nat Nanotechnol, 2012,7(7):459-64

(12) Li L, Fang CJ, Ryan JC, et al. Binding and uptake of H-ferritin are mediated by human transferrin receptor-1. Proc Natl Acad Sci U S A, 2010,107(8):3505-10

Version 2

Abstract

Abstract

Stem cell niche is critical for regulating the behavior of stem cells. Drosophila neural stem cells (Neuroblasts, NBs) are encased by glial niche cells closely, but it still remains unclear whether glial niche cells can regulate the self-renewal and differentiation of NBs. Here we show that ferritin produced by glia, cooperates with Zip13 to transport iron into NBs for the energy production, which is essential to the self-renewal and proliferation of NBs. The knockdown of glial ferritin encoding genes causes energy shortage in NBs via downregulating aconitase activity and NAD+ level, which leads to the low proliferation and premature differentiation of NBs mediated by Prospero entering nuclei. More importantly, ferritin is a potential target for tumor suppression. In addition, the level of glial ferritin production is affected by the status of NBs, establishing a bicellular iron homeostasis. In this study, we demonstrate that glial cells are indispensable to maintain the self-renewal of NBs, unveiling a novel role of the NB glial niche during brain development.

Introduction

Neural stem cells (NSCs) possess the remarkable ability to self-renew for maintaining their identity and differentiate into diverse neural cell types required for neurogenesis during brain development. The self-renewal, differentiation, and proliferation of NSCs are regulated by both intrinsic programs inside stem cells and extrinsic cues from the surrounding niches. The mammalian NSC niche found in the neurogenic regions is composed of diverse cell types including choroid plexus, vascular endothelia cells, pericytes, microglia, and meninges, which regulate NSC behaviors such as proliferation, self-renewal, and differentiation (Bjornsson, Apostolopoulou, Tian, & Temple, 2015). The diversity and complexity of these cell populations confer niche the ability of integrating local and systemic signals, but make it complicated to study the interaction between niches and NSCs in mammals, which is a critical issue to understand neurogenesis and improve clinical applications of NSCs.

Drosophila larval central nervous system (CNS), composed of central brain (CB), optic lobe (OL) and ventral nerve cord (VNC), is a simple but sophisticated model to decipher the regulatory mechanisms underlying NSC behaviors mediated by niche signals. In this system, Drosophila NSCs, called neuroblasts (NBs), and their progenies are encased by a variety of glial cells including perineurial glia, subperineurial glia, cortex glia, astrocyte-like glia and ensheathing glia (Hartenstein, 2011; Ito, Urban, & Technau, 1995; Pereanu, Shy, & Hartenstein, 2005). These glial cells not only form the physical support structure for NBs, but also provide extrinsic cues that regulate NB reactivation, survival, and proliferation (Read, 2018; Speder & Brand, 2014). However, it remains elusive whether the glial-cell-forming niche is involved in the regulation of the prominent characteristics of NBs, self-renewal and differentiation.

Iron is a vital micronutrient for almost all living organisms since it participates in many crucial biological processes, including mitochondrial electron transport and cellular respiration that both support ATP production used for most cellular activities. High levels of iron are found in many organs including the brain, where iron is required for myelination and formation of neuronal dendritic trees (Lozoff et al., 2006; Moller et al., 2019; Rice & Barone, 2000). Iron deficiency in infants and early childhood has been well documented to be associated with long-term cognitive, social, and emotional problems (Fretham, Carlson, & Georgieff, 2011). Thus, it is intriguing to investigate whether iron is also required in NSCs during neurogenesis.

Ferritin, a ubiquitous protein present in most organisms, is composed of 24 subunits that assemble into a hollow-sphere complex capable of storing up to 4,500 iron atoms in a bioavailable form (Andrews, 2005; Knovich, Storey, Coffman, Torti, & Torti, 2009). Ferritin in both mammal and insect is a heteropolymer, consisting of heavy and light chain homologue subunits. In Drosophila, these subunits are encoded by ferritin 1 heavy chain homologue (Fer1HCH) and ferritin 2 light chain homologue (Fer2LCH) respectively (Arosio, Ingrassia, & Cavadini, 2009; Pham & Winzerling, 2010), which both contain secretion signal peptides that direct them to the ER during translation (Lind, Ekengren, Melefors, & Soderhall, 1998). Consequently, Drosophila ferritin is confined to the secretory pathway and abundant in the hemolymph (Nichol, Law, & Winzerling, 2002). Ferritin or iron manipulation in Drosophila glia regulate the adult behavior, such as locomotion (Kosmidis et al., 2011; Navarro, Botella, Metzendorf, Lind, & Schneuwly, 2015). Previous researches have shown that Drosophila ferritin not only directly participates in dietary iron absorption by removing iron from enterocytes across the basolateral membrane, but competes with transferrin to deliver iron between the gut and the fat body (Tang & Zhou, 2013; Xiao, Liu, Zhao, Wang, & Zhou, 2019). These findings underscore the pivotal role of ferritin in iron transport. Interestingly, iron-loaded ferritin, serving as an essential mitogen, promotes proliferation of cultured Drosophila cells (Li, 2010). Therefore, it is tempting to investigate whether ferritin is required for self-renewal and proliferation of NSCs during brain development.

Here we show that glial ferritin supplies iron into NBs to form iron-sulfur (Fe-S) clusters, which is necessary for the production of ATP, providing the energy required for self-renewal and proliferation of NBs, highlighting ferritin as a potential target for suppressing tumor.

Results

Glial ferritin is required for NB maintenance and proliferation

To identify new niche factors involved in the balance between NB self-renewal and differentiation, we used repo-GAL4 to perform an RNAi mediated screen in glial cells. Fer1HCH and Fer2LCH, two ferritin subunit encoding genes, were identified to be essential for NB maintenance and brain development. Specifically, knocking down Fer1HCH or Fer2LCH in glial cells significantly reduced the number of NBs in the CB and VNC at the third-instar larval stage, which resulted in decreased size of the CNS (Figure 1A and B and Figure 1-figure supplement 1A). To further assess the impact on NB proliferation, we employed the mitotic marker phosphohistone H3 (PH3) and the thymidine analogue 5-ethynyl-2’-deoxyuridine (EdU), revealing a substantial decrease in the proliferation rate of NBs upon glial ferritin knockdown (Figure 1A and C and Figure 1-figure supplement 1B and C). These results indicated that glial ferritin is required for maintaining NBs and their proliferation.

Ferritin in glia is required for NB maintenance and proliferation

(A) NB number and proliferation rate after Fer1HCH or Fer2LCH RNAi in glia.

(B) Quantification of NB number in CB (A).

(C) Quantification of NB proliferation rate in CB (A).

(D and E) RT-PCR analysis of Fer1HCH and Fer2LCH mRNA levels (D) and normalized quantification (E) after glial ferritin knockdown.

(Fand H) Western blot of Fer1HCH (F) and Fer2LCH (H) after ferritin knockdown in glia. Tubulin was used as a loading control.

(G and I) Quantification of normalized Fer1HCH intensity in (G) and Fer2LCH intensity in (I).

(J) Rescue of NB number and proliferation when simultaneously overexpressing RNAi-resistant Fer1HCH and Fer1HCH RNAi (or Fer2LCH and Fer2LCH RNAi) in glia.

(K and L) Quantification of NB number and proliferation rate in (J).

(B, n=11; C, n=10; E, G and I, n=3; K, n=8,8,9,7,9; L, n=9,8,8,7,12; ns, not significant; *: p<0.05; **: p<0.01; ***: p<0.001; ****: p<0.0001; Dpn, NB nuclei, red; PH3, green; DAPI, nuclei, blue).

Glial ferritin knockdown leads to low NB proliferation and number

(A) Quantification of NB number in VNC.

(B) Quantification of NB proliferation rate in VNC.

(C) EdU incorporation in CNS after knocking down glial ferritin.

(D) Knocking down Fer2LCH by nrv2-GAL4 led to similar but weaker phenotype in brain in comparison with pan-glial ferritin knockdown.

(A, n=11; B, n=10; ****: p<0.0001)

Ferritin knockdown in different glial subpopulations did not induce any discernible phenotype

(A) Knocking down ferritin genes in subperineurial glia, perineurial glia, ensheathing glia and astrocyte-like glia, respectively

Verify the phenotype induced by glial ferritin knockdown using different manipulations

(A) NB number and proliferation in other RNAi lines targeting on ferritin.

(B) Rescue of NB number in VNC when simultaneously overexpressed RNAi-resistant Fer1HCH and Fer1HCH RNAi (or Fer2LCH and Fer2LCH RNAi) in glia.

(C) Rescue of NB proliferation rate in VNC when simultaneously overexpressed RNAi-resistant Fer1HCH and Fer1HCH RNAi (or Fer2LCH and Fer2LCH RNAi) in glia.

(D) NB number and proliferation in Fer1HCHDN. (B, n=8,7,9,7,8; C, n=9,8,8,7,12; ****: p<0.0001)

To identify the glial subpopulation in which ferritin functions, we employed subtype-restricted GAL4 to perform ferritin knockdowns (Table 1). We observed that knocking down ferritin genes in cortex glia led to a similar yet milder phenotype compared to that observed in the pan-glia knockdown (Figure 1-figure supplement 1D). This milder effect may be attributed to potential compensatory mechanisms from other glial subpopulations. Interestingly, knocking down ferritin in other glial subtypes did not yield any discernible phenotype (Figure 1—figure supplement 2). These findings suggest that ferritin functions collectively across all glial populations to regulate the activity of NBs.

We confirmed the phenotype with multiple independent RNAi lines for Fer1HCH or Fer2LCH, and observed similar abnormalities of larval NBs and brain development (Figure 1-figure supplement 3A). Hereafter, UAS-Fer1HCH-RNAi (THU5585) and UAS-Fer2LCH-RNAi (TH01861.N) were used for most studies, and we only quantified NB number and proliferation rate in the CB for convenience in this research.

To evaluate the efficiency of RNAi for Fer1HCH or Fer2LCH, RT-PCR was performed with total RNA isolated from whole brain dissected from third-instar larvae. The abundance of Fer1HCH or Fer2LCH mRNA was significantly reduced in glial Fer1HCH or Fer2LCH knockdown respectively (Figure 1D and E). Western blot analysis of larval brain protein extract using polyclonal antibodies against Fer1HCH or Fer2LCH further confirmed that both subunits were down-regulated in glial knockdown of either subunit (Figure 1F-I), consistent with previous research in Drosophila (Missirlis et al., 2007). To exclude the possibility of RNAi off-target effects, we used repo-GAL4 to overexpress RNAi-resistant Fer1HCH (or Fer2LCH) containing silent mutations in the region targeted by shRNA, which resulted in a significant increase in the number and proliferation rate of NBs compared to glial ferritin knockdown alone (Figure 1J-L and Figure 1-figure supplement 3B and C). Additionally, similar but weaker phenotypes were observed when Fer1HCHDN, a ferroxidase-inactive form of Fer1HCH (Missirlis et al., 2007), was overexpressed in glial cells (Figure 1-figure supplement 3D). These results support the conclusion that ferritin in glial cells is essential for the maintenance and proliferation of NBs, suggesting that iron homeostasis is involved in this process.

Ferritin is produced mainly in glial cells and secreted into NBs through a vesicle-dependent pathway

To investigate the role of iron homeostasis in Drosophila CNS, we tried to knock down genes involved in iron transport, storage, and regulation in NBs, neurons, and glial cells, respectively. Interestingly, only glial knockdown of ferritin-related genes led to NB loss and defective brain development. It appears that ferritin generated in glial cells is the only crucial protein for regulating the behavior of NBs and, ultimately, the larval brain development. To recapitulate the endogenous ferritin expression pattern, we took advantage of a Fer1HCH/Fer2LCH enhancer trapped GAL4 line to drive a variant of DsRed protein with a nuclear localization signal (DsRed Stinger). Surprisingly, our immunostaining experiments revealed that most, if not all, DsRed was co-localized with the glia nuclear protein Repo, instead of the NB nuclear protein Dpn (Figure 2A and A’). We also used another Fer2LCH enhancer trapped GAL4 line to verify this result, which was consistent with above result (Figure 2—figure supplement 1). These results further confirm that glial cells are the main producers of ferritin in the larval brain.

Ferritin is produced mainly in glial cells in the Drosophila CNS and secreted into NBs through a vesicle dependent pathway

(A) The pattern of DsRed Stinger driven by Fer1HCH/Fer2LCH-GAL4 in CNS. (A’) The magnification of (A).

(B) The distribution of ferritin labeled by Fer1HCHG188 in CNS of control and glial ferritin knockdown.

(B’) The magnification of left panel in (B).

(C) Overexpressed ferritin tagged with GFP and mCherry in glia secreted into NBs.

(D) The distribution of Fer1HCHG188 when blocking glial vesicular trafficking via overexpressing dominant-negative dynamin.

(D’) The magnification of (D).

(E) Quantification of (D), GFP signal in NBs when overexpressing shi ts in glia.

(All white circles indicate NB position and white arrows point glial cell; E, n=50, 48; **: p<0.01)

Fer2LCH-GAL4 was used to validate the pattern of Fer1HCH/Fer2LCH-GAL4

(A) Most DsRed driven by Fer2LCH-GAL4 was colocalized with glial marker Repo.

Fer1HCHG188 encodes a GFP-tagged Fer1HCH subunit (Missirlis et al., 2007), which labels the distribution of Fer1HCH, representing the pattern of ferritin. As shown in the Figure 2B and B’, ferritin was detected in NBs besides glial cells, despite that ferritin is specifically produced in glial cells. To explore the possibility of ferritin transfer from glial cells to NBs, we observed the GFP signal of Fer1HCHG188 with glial ferritin knockdown, and found that GFP in the brain was almost completely depleted (Figure 2B). Furthermore, we used repo-GAL4 to drive the expression of GFP-tagged Fer1HCH and mCherry-tagged Fer2LCH in glial cells. We observed that both Fer1HCH::GFP and Fer2LCH::mCherry were present in NBs, while the control GFP and mCherry were restricted to glial cells (Figure 2C). Altogether, we conclude that ferritin is generated in glial cells and then transferred to NBs in the larval brain.

In Drosophila, both ferritin subunits contain secretion signal peptides that direct protein to the endoplasmic reticulum during translation, which allows the ferritin to be transported out of the cell. To investigate whether the ferritin was transported from glia to NBs via the vesicular transport, we expressed a temperature-sensitive dominant-negative Shibire dynamin protein (shits) to block vesicle trafficking from glial cells, and found that Fer1HCH::GFP accumulated in glia, accompanying with significantly diminished Fer1HCH signal in NBs (Figure 2D, D’ and E), suggesting glial ferritin secretes into NBs through vesicle system.

Glial ferritin defects lead to iron deficiency in NBs

Given that ferritin is required for iron storage as well as delivery, glial ferritin knockdown might increase free iron level, leading to glial iron overload. Alternatively, it will block the iron transport from glia to NBs, resulting in NB iron deficiency. To determine the underlying cause of the brain development defects observed upon glial ferritin knockdown, we used iron chelator or iron salt ferric ammonium citrate (FAC) to manipulate iron intake and detect how it will influence NBs. We observed that depriving iron intake by adding iron chelator bathophenanthrolinedisulfonic acid disodium (BPS) to the food led to a dramatic decrease in the number and proliferation rate of NBs, compared to control larvae on normal food (Figure 3A). To investigate the effect of feeding behavior in different food, we tested the feeding behavior of fly by incorporating 1% Brilliant Blue (sigma, 861146) into different food (Tanimura, Isono, Takamura, & Shimada, 1982) and found the amount of dye in the fly body was similar between normal group and BPS group (Figure 3—figure supplement 1A), which suggested that BPS almost did not affect the feeding behavior. In line with this result, blocking iron absorption by knocking down intestinal ferritin (Tang & Zhou, 2013) resulted in the similar defects of brain (Figure 3-figure supplement 1B-D). However, supplementing the diet with FAC did not induce any observable defects (Figure 3A). We also tested another iron chelator deferiprone (DFP) (Soriano et al., 2013), and found that DFP alone did not cause any brain defects in wild-type (WT) larvae, but it significantly exacerbated the effect of glial ferritin knockdown on defective NB maintenance (Figure 3B and C). Furthermore, we attempted to rescue the brain defects observed in glial ferritin knockdown by providing dietary FAC, but the effect was negligible (Figure 3-figure supplement 1E and F). This may be due to the possibility that the knockdown of glial ferritin largely, if not completely, prevented the transport of iron from glia to NBs, which hindered the additional iron from entering NBs. These results indicate that the NB loss phenotype induced by glial ferritin knockdown is primarily due to iron deficiency in NBs.

Glial ferritin defects lead to iron deficiency in NBs

(A) The conditions of brain development in Drosophila with feeding iron chelator BPS or supplementary iron FAC.

(B) Iron depletion by adding DFP to food exacerbated the phenotype of NB loss induced by glial ferritin knockdown.

(C) Quantification of NB number in (B).

(D) Zip13 in CNS is produced mainly by glial cells, but not NBs.

(D’) The magnification of (D). (All white circles indicate NB position and white arrows point glial cells)

(E) Brain defects in Zip13 knockdown or double knockdown of Zip13 and Fer2LCH. (F and G) Quantification of NB number and proliferation rate in (E).

(H) Iron supplement can rescue the NB proliferation in Zip13 RNAi driven by repo-GAL4.

(I) Quantification of NB proliferation in (H).

(C, n=6; F, n=6,10,11,8; G, n=13,10,10,5; I, n=8; ns, not significant; *: p<0.05; **: p<0.01; ****: p<0.0001)

Brain defects induced by glial ferritin knockdown were caused by iron deficiency

(A) Test the feeding behavior via adding Brilliant Blue to normal food or BPS food.

(B) Iron deficiency by blocking iron absorption in gut also induced similar brain defects caused by BPS addition.

(C) Quantification of NB number in (B).

(D) Quantification of NB proliferation in (B).

(E) Brain development of glial ferritin knockdown after feeding FAC.

(F) The ratio of brain with NB loss after adding FAC.

(G) NB number and proliferation of Zip13 mutant.

(H and I) Quantification of NB number and proliferation rate in (G).

(C, n=10; D, n=5; F, n=3; H, n=10,9; I, n=9,8; *: p<0.05; **: p<0.01; ***: p<0.001; ****: p<0.0001)

ROS accumulates in CNS after glial ferritin knockdown, but inhibiting ROS cannot restore NB number and proliferation

(A) ROS level marked by Gstd-GFP after knocking down ferritin in glia.

(B) Quantification of ROS level in (A).

(C) Overexpress antioxidant sod1 or cat in glial cells.

(D) Overexpress antioxidant sod1 or cat in NBs.

(E) Mitochondrial morphology by TEM upon glial ferritin knockdown.

(F) Quantification of normal mitochondria in (E). Intact mitochondria with their cristae visible were considered normal.

(G) Western blot of 4-HNE after ferritin knockdown in glia. Actin was used as a loading control.

(H) Quantification of normalized 4-HNE intensity in (G).

(B, n=5,4,5; F, n=23,27,28; H, n=3; ****: p<0.0001; ns, not significant)

Glial ferritin knockdown does not induce apoptosis in the CNS

(A) Apoptotic signal labeled by Caspase-3 of glial ferritin knockdown.

(B) Quantification of Caspase-3 negative NB ratio.

(C) TUNEL staining of Fer1HCH RNAi driven by repo-GAL4.

(D) Quantification of TUNEL positive cell per 1×105 μm2.

(E) P35 overexpression in glia could not rescue brain defects of ferritin knockdown.

(F) Glial morphology was intact when knocking down glial ferritin. (B, n=5,4,4; D, n=3; ns, not significant)

Zip13, generally known as a zinc transporter, functions as an iron transporter, cooperating with ferritin on iron transport in Drosophila intestine (Xiao, Wan, Fan, Tang, & Zhou, 2014). Specifically, knockdown of Zip13 leads to failure of iron loading in secretory vesicles and consequently defective iron efflux (Xiao et al., 2014). To investigate whether the NB loss phenotype of glial ferritin knockdown is caused by defective iron transport, we next sought to identify the role of Zip13 in the larval brain. Firstly, we examined the expression pattern of Zip13 and found that, similar to ferritin, the DsRed stinger driven by Zip13-GAL4 was co-localized with Repo in glial cells rather than Dpn in NBs (Figure 3D and D’). Subsequently, we knocked down Zip13 in glial cells and observed a similar but weaker phenotype with the slightly reduced NB number and proliferation rate (Figure 3E-G). Consistently, the defective brain development was confirmed in Zip13 mutant (Figure 3-figure supplement 1G-I). Moreover, the effect of double knockdown of both Zip13 and Fer2LCH was much more severe than that of either single knockdown, with a significant reduction in NB number and proliferation rate (Figure 3E-G). Furthermore, we found that the reduced proliferation rate caused by Zip13 knockdown could be partially rescued by iron supplementation in the food (Figure 3H and I). Overall, these results suggest that glial ferritin collaborates with Zip13 to deliver iron from glia to NBs.

Since previous studies have demonstrated the essential role of glial cells in NB survival and proliferation (Bailey et al., 2015; Read, 2018), it is possible that iron overload in glial cells up-regulated the level of reactive oxygen species (ROS), leading to apoptosis of glial cells, which ultimately resulted in the loss of NBs. To rule out this possibility, firstly, we examined the ROS level of glial ferritin knockdown via a transgenic reporter GstD-GFP for oxidative stress signaling (Sykiotis & Bohmann, 2008). GstD1 is a prototypical oxidative stress response gene (Sawicki, Singh, Mondal, Benes, & Zimniak, 2003). We found an obvious increase as expected (Figure 3-figure supplement 2A and B). However, overexpressing anti-oxidant genes in glial cells or NBs did not restore the NB number and proliferation (Figure 3-figure supplement 2C and D). ROS accumulation leads to lipid peroxidation, which in turn results in ferroptosis (Hirschhorn & Stockwell, 2019). Mitochondria in cells undergoing ferroptosis typically exhibit a characteristic shrinkage and damage, while the nuclei remain structurally intact (Mumbauer, Pascual, Kolotuev, & Hamaratoglu, 2019). In order to investigate whether the NB loss is due to ferroptosis, we examined the mitochondrial morphology using Transmission Electron Microscopy (TEM). However, our result showed that the mitochondrial double membrane and cristae were clearly visible whether in the control group or glial ferritin knockdown group, which suggested that ferroptosis was not the main cause of NB loss upon glial ferritin knockdown (Figure 3-figure supplement 2E and F). In addition, we investigated the level of 4-HNE that indicates lipid peroxidation, the other hallmark of ferroptosis. This result showed that 4-HNE level did not change significantly after ferritin knockdown (Figure 3-figure supplement 2G and H), suggesting that lipid peroxidation was stable, which supported the exclusion of the ferroptosis in glial ferritin knockdown. Taken with the findings from iron chelator feeding together, we conclude that NB loss resulting from glial ferritin knockdown is not a consequence of ferroptosis. Secondly, we examined the apoptosis in larval brain cells via Caspase-3 or TUNEL staining, and found the apoptotic signal remained unchanged after glial ferritin knockdown (Figure 3-figure supplement 3A-D). Furthermore, we overexpressed the baculovirus protein P35 to inhibit apoptosis in glial cells, but failed to restore the lost NBs (Figure 3-figure supplement 3E). Given that glial morphology is important for NB lineages (Speder & Brand, 2018), we used GFP driven by repo-GAL4 to denote glia cells and found that glial chamber was almost intact after glial ferritin knockdown (Figure 3-figure supplement 3F). Together, we conclude that it is iron deficiency in NBs, rather than iron overload in glial cells, that leads to NB loss induced by ferritin defects.

Glial ferritin defects result in impaired Fe-S cluster activity and ATP production

To further prove that defective iron transport leads to the iron deficiency in NBs, we examined the iron availability by measuring aconitase activity in the brain. The activity of aconitase is sensitive to the level of Fe-S synthesis, so it can serve as an indicator for the availability of iron in the cell (Haile et al., 1992; Tong & Rouault, 2006). Aconitase activity is determined in a coupled enzyme reaction in which citrate is converted to isocitrate by aconitase (Aconitase Activity Assay Kit, sigma, MAK051). As shown in the Figure 4A, aconitase activity of ferritin knockdown diminished significantly, indicating that the availability of iron in the brain was reduced. This result also implies Fe-S clusters in NBs are probably deficient when knocking down ferritin in glia. Fe-S clusters are essential to activate the aconitase that converts citrate into isocitrate in the tricarboxylic acid cycle (TCA cycle) in mitochondria, and are one type of critical electron carrier for electron transport chain (ETC), especially for the Complex I, II and III function. Considering that TCA cycle and ETC are critical biological processes for ATP production, we proposed that knockdown of glial ferritin might lead to ATP deficiency due to the block of TCA cycle and ETC, which resulted in the failure of maintaining NBs and their proliferation. To verify this hypothesis, we first assessed the function of Fe-S clusters in NBs by knocking down mitochondrial Fe-S protein Nfs1 and the components of the cytosolic iron-sulfur protein assembly (CIA) complex. After knocking down Nfs1 and the CIA component CG17683, galla2, and ciao1 in NBs respectively, the proliferation level of NBs decreased significantly and brain size became smaller (Figure 4B and C), which phenocopied the knockdown of Fer1HCH or Fer2LCH in glia. This result implies that Fe-S clusters are indispensable for NBs to maintain their proliferation.

Glial ferritin defects result in impaired Fe-S cluster activity and ATP production

(A) The determination of cytosolic aconitase activity.

(B) Knockdown of mitochondrial iron-sulfur protein Nfs1 and CIA complex in NBs.

(C) Quantification of NB proliferation in (B).

(D) The schematic diagram of NB sorting procedure.

(E) Venn diagram of downregulated genes after glial ferritin knockdown.

(F) GO enrichment of down-regulated genes in glial ferritin knockdown comparing with control.

(G) KEGG pathway enrichment of down-regulated genes after glial ferritin knockdown.

(H) NAD+ level was indicated by SoNar.

(I) Quantification of normalized SoNar signal in (H).

(J) The determination of ATP level in CNS.

(K) NB number in glial ferritin knockdown was rescued by Ndi1 overexpression in NBs.

(L) Quantification of NB number in (K).

(A and J, n=3; C, n=5,7,6,6,9; I, n=37,42; L, n=7,7,6,6,7; *: p<0.05; **: p<0.01; ***: p<0.001; ****: p<0.0001)

Validation of RNA-seq data by qRT-PCR using sorted NB lineages and Ndi1 overexpression

(A) qRT-PCR analysis of ND-15 and CG15715 to validate the enrichment result.

(B) Ndi1 overexpression could restore ATP production. (A and B, n=3; *: p<0.05; **: p<0.01)

To identify the biological processes related to Fe-S clusters in NB lineages, we performed transcriptome analysis of sorted NB lineages from glial ferritin knockdown as well as the control groups. We used NB driver Insc-GAL4 to express UAS-DsRed, labeling NB lineages, and then sorted these cells via fluorescence-activated cell sorting (FACS) (Figure 4D) (Harzer, Berger, Conder, Schmauss, & Knoblich, 2013). We found that thousands of genes were significantly altered in the Fer1HCH and Fer2LCH knockdown groups compared to the control group. Specifically, 880 genes were simultaneously downregulated among the significantly changed genes in both the Fer1HCH and Fer2LCH knockdown groups contrasted to the control group (Figure 4E). Given that iron deficiency in NBs probably leads to the defects of energy production and biosynthesis, we performed gene enrichment analysis in these genes based on Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG). GO term analysis showed that mitochondrial organization, translation, transport, oxidative phosphorylation (OxPhos), and ETC were enriched in the Biological Processes (Figure 4F), indicating NB mitochondrial functions including OxPhos and ETC were abnormal after glial ferritin knockdown. Moreover, NADH dehydrogenase and ribosome related terms were also enriched whether in the Cellular Component or Molecular Function (Figure 4F), suggesting that NADH dehydrogenase and biosynthesis were defective. Further, we validated the enriched term associated with mitochondrial function by qRT-PCR (Figure 4-figure supplement 1A). And the enriched KEGG pathways included OxPhos and ribosome (Figure 4G), aligning with the GO enrichment results and supporting the hypothesis that glial ferritin knockdown disrupted mitochondrial function including OxPhos and ETC as well as further biosynthesis.

To confirm whether the function of NADH dehydrogenase was disrupted, we used the SoNar that could bind to NAD+ with specific conformation and excitatory wavelength to monitor the NAD+ level (Bonnay et al., 2020; Zhao et al., 2016). SoNar was driven by the Insc-GAL4 to indicate the NAD+ level in NBs. Considering the possibility of protein accumulation in NBs due to the lower proliferation rate in glial ferritin knockdown group, the signal of SoNar was normalized to DsRed Stinger driven by Insc-GAL4. We observed significantly declined SoNar/DsRed after knocking down Fer2LCH (Figure 4H and I), suggesting the normalized level of NAD+ in NBs was decreased, which supports the conclusion that the activity of NADH dehydrogenase was declined. To further verify the change of energy production, we utilized the ATP Determination Kit which offers a convenient bioluminescence assay for quantitative determination of ATP with recombinant firefly luciferase and its substrate D-luciferin. We found that ATP level decreased significantly after glial ferritin knockdown when compared to the control (Figure 4J). Therefore, we tried to supply the energy for NBs by restoring bioenergy-dependent NAD+ metabolism. To validate this, we firstly determined the ATP production after overexpressing Ndi1 in NBs upon glial Fer2LCH knockdown. The data suggested that expression of Ndi1 can restore ATP production (Figure 4-figure supplement 1B). Furthermore, the NB number could be partially rescued by overexpressing proton-pump-independent yeast mitochondrial NADH dehydrogenase Ndi1 in NBs under the background of glial ferritin knockdown (Figure 4K and L) (Bonnay et al., 2020). In summary, the iron deficiency leads to NB loss due to the insufficient energy required for essential biological processes.

Glial ferritin maintains NBs via preventing premature differentiation

To investigate how glial ferritin knockdown leads to NB loss, we firstly examined whether NB origin was affected via altering the timing of ferritin knockdown. We used the temperature-sensitive GAL4 inhibitor GAL80ts to suppress ferritin RNAi at a permissive temperature (18℃) during embryonic stage, and then shifted to a restrictive temperature (29℃), or vice versa until we dissected them at the third-instar larval stage. Ferritin knockdown during the larval stage led to severe brain development defects (Figure 5A and B), while the knockdown only during the embryonic stage showed no apparent abnormality (Figure 5-figure supplement 1A and B). Together, these results indicate that glial ferritin is required for NB maintenance at the larval stage. Consistent with this result, we found no obvious defects of brain at the first-instar larval stage when knocking down glial ferritin (Figure 5-figure supplement 1C and D). Since newborn NBs are delaminated from the neuroectoderm during the embryonic stage (stage 9 to 11), and ferritin RNAi only during the embryonic stage did not show any phenotype, suggesting that NB loss after glial ferritin knockdown is not ascribed to defective NB origin.

Glial ferritin defects lead to the premature differentiation of NBs

(A) Temporal window of glial ferritin maintaining NBs. The temperature shift of flies with glial ferritin knockdown from the permissive temperature (18℃) to the restrictive temperature (29℃).

(B) Quantification of NB number and proliferation in (A).

(C) Pros staining of knocking down glial ferritin.

(D) Quantification of ratio of Pros+Dpn+ NBs to Dpn+ NBs.

(E) Knockdown of pros in NBs rescues the NB number.

(F) The ratio of brain with NB loss in (E).

(B, n=6, 5, 6; n=6, 5, 4; D, n=5; F, n=3; *: p<0.05; ***: p<0.001; ****: p<0.0001)

NB loss upon glial ferritin knockdown is not due to NB origin and apoptosis

(A) Temporal window of glial ferritin maintaining NBs. The temperature shift of flies with glial ferritin knockdown from the restrictive temperature (29℃) to the permissive temperature (18℃).

(B) Quantification of NB proliferation in (A).

(C) The CNS of the first-instar larvae with glial ferritin knockdown.

(D) Quantification of NB number and proliferation in (C).

(E) The CNS of the second-instar larvae with glial ferritin knockdown.

(F) Quantification of NB number and proliferation in (E).

(G) P35 overexpression in NBs could not rescue the brain defects induced by glial ferritin knockdown.

(B, n=5; D, n=6; 8,9,8; F, n=11,9,9; 8,9,8; *: p<0.05; **: p<0.01; ns, not significant)

Ndi1 overexpression can restore Pros localization

(A) Pros staining of Ndi1 overexpression upon glial ferritin knockdown.

(B) Quantification of ratio of Pros+Dpn+ NBs to Dpn+ NBs in (A). (B, n=6,9; ***: p<0.001)

Glial ferritin defects do not affect NB reactivation and asymmetric division

(A) aPKC localization in NB after knocking down Fer2LCH.

(B) NB size upon glial ferritin knockdown was not changed.

(C) Activation of Insulin signaling could not restore NB number and proliferation.

(D) Quantification of NB proliferation in (B). (B, n=100; D, n=8, 5; ns, not significant)

Further, we explored the earliest stage when the phenotype (NB number and proliferation) existed after glial ferritin knockdown. The result showed that NB proliferation decreased significantly, but NB number declined slightly at the second-instar larval stage (Figure 5—figure supplement 1E and F), suggesting that brain defect of glial ferritin knockdown manifests at the second-instar larval stage.

As mentioned before, we did not detect any apoptotic signal in the larval brain after glial ferritin knockdown, suggesting that there is no association between apoptosis and NB loss. The double binary systems, lexA/lexAop and GAL4/UAS, were employed to further exclude the possibility of apoptosis. We overexpressed P35 in NBs by Insc-GAL4 and knocked down ferritin by repo-lexA, which failed to restore the decreased number of NBs (Figure 5-figure supplement 1G). These results reinforce the notion that NB loss occurs independent of the apoptotic pathway.

Next, we hypothesized that the NB loss might be attributed to nuclear-Prospero (Pros)-dependent premature differentiation of NBs (Cabernard & Doe, 2009; Choksi et al., 2006). Pros in NB nuclear was hardly detected in third-instar WT larvae, but was increased significantly with glial ferritin knockdown (Figure 5C and D). To further verify that the NB loss caused by glial ferritin knockdown was due to nuclear-Pros-dependent premature differentiation, we used the Insc-GAL4 to drive UAS-pros-RNAi in order to prevent the NB loss in glial ferritin knockdown background. We found that the NB number was substantially restored (Figure 5E and F). Interestingly, despite of dramatic NB number increase, the decreased proliferation rate was not rescued. We proposed that pros knockdown in NBs blocked premature terminal differentiation, but it could not replenish the reduced energy caused by glial ferritin knockdown. As NB self-renewal and proliferation are both energy-consuming processes, the self-renewal was rescued via inhibiting differentiation by Insc-GAL4 driving UAS-pros-RNAi, but at the expense of energy consumption for proliferation, leaving the proliferation rate unchanged, or even lower. Together, we conclude that glial ferritin maintains the number of NBs by preventing the nuclear-Pros-dependent premature differentiation.

Furthermore, we also investigated whether overexpression of Ndi1 could restore Pros localization in NBs. This result showed that overexpressing Ndi1 could significantly restore Pros localization in NBs (Figure 5-figure supplement 2), which supported that energy shortage induced the entry of Pros into nuclei, leading to the premature differentiation of NBs.

Considering that asymmetric division defects in NBs may lead to premature differentiation, we also explored the asymmetric division by staining the classical asymmetric marker aPKC and found it displayed a crescent at the apical cortex based on the daughter cell position whether in control or glial ferritin knockdown (Figure 5-figure supplement 3A). This result indicated that there was no obvious asymmetric defect after glial ferritin knockdown.

Given that reduced proliferation is a distinctive characteristic of quiescent NBs, we tried to determine whether NBs were in a quiescent state after glial ferritin knockdown by examining the other two key features associated with quiescent NBs: decreased cell size and the formation of the extended cellular protrusions (Chell & Brand, 2010; Ly et al., 2019). However, we found that the NB size remained at the same level as control, and the extended cellular protrusion was not observed after glial ferritin knockdown (Figure 2B’ and Figure 5-figure supplement 3B). Furthermore, since insulin signaling is necessary and sufficient to reactivate NBs from quiescence (Sousa-Nunes, Yee, & Gould, 2011), we tried to overexpress a constitutively active form of InR (InRact) in NBs to promote reactivation at the background of glial ferritin knockdown. However, InRact overexpression failed to ameliorate the NB abnormalities including the decreased proliferation (Figure 5-figure supplement 3C and D). In conclusion, NBs exhibited a low proliferation rate after glial ferritin knockdown, but were not in a quiescent state.

Ferritin functions as a potential target for tumor suppression

We have demonstrated that glial ferritin is essential for NB self-renewal and proliferation by providing glia-to-NB iron transport to support energy production in NBs. We next tried to determine whether manipulating this biological function could suppress tumor development. We induced tumor development by driving UAS-brat-RNAi or UAS-numb-RNAi under the control of Insc-GAL4, and knocked down ferritin genes under the control of repo-lexA. We observed that the size of tumors was strikingly decreased in the larval brains lacking glial ferritin (Figure 6A and B and Figure 6-figure supplement 1A and B). This result suggests that glial ferritin is required for tumor development, highlighting a potential target for tumor treatment.

Ferritin functions as a potential target for tumor suppression

(A) Knockdown of ferritin in glia could inhibit the tumor induced by numb RNAi.

(B) Quantification of larval brain size in (A).

(C) Iron chelator BPS supplemented in the food suppressed brain tumor growth.

(D) Quantification of larval brain size in (C).

(E) Bioluminescence signal in mice with glioma on day 14.

(F) Quantification of bioluminescence signal in (E).

(G) Kaplan-Meier survival curve of mice treated with DFP.

(B, n=6,7,5; D, n=5,6,6; F, n=6,7; G, n=7,6; *: p<0.05; **: p<0.01 ***: p<0.001; ****: p<0.0001)

Iron chelator inhibits tumor progression in Drosophila and mice

(A) Knockdown of ferritin in glia could inhibit the tumor induced by brat RNAi.

(B) Quantification of larval brain size in (A).

(C) Iron chelator BPS suppressed brain tumor growth of brat RNAi.

(D) Quantification of larval brain size in (C).

(E) The appearance and HE staining of brains in mice with DFP injection on day 7, 14 and 21. Scar bar: 1000 μm.

(F and G) Quantification of HE staining in murine brains on day 14 (B) and 21 (C). (B, n=10,7; D, n=8,6; F, n=6; G, n=9,8; ***: p<0.001; ****: p<0.0001.)

To verify whether iron is involved in the tumor suppression induced by blocking glial ferritin, we tried to decrease iron level by adding the iron chelator BPS in the Drosophila food. Tumor formation in the brain at the third instar larval stage was induced with the same method as mentioned above. The addition of BPS into dietary food significantly reduced the tumor size (Figure 6C and D and Figure 6-figure supplement 1C and D), suggesting that iron is required for tumor development in Drosophila larval brain. Considering the conservation of iron’s fundamental function as an essential nutrient for cell growth and proliferation, we thus tried to suppress brain tumor in mice via iron deficiency. Initially, we constructed orthotopic glioma model in C57BL/6J mice. One week after inoculation of glioma cells (GL261-luc), the iron chelator DFP was intraperitoneally injected every two days based on the murine weight until being dissected on day 7, 14, and 21, respectively. Histological examination with HE staining revealed that the size of glioma in mice, injected with DFP, decreased significantly on day 14 and 21 (Figure 6-figure supplement 1E-G). Since glioma cells GL261-luc carry the luciferase gene, we injected the substrate luciferin potassium salt into the tumor-bearing mice and detected the bioluminescence signal of mouse gliomas using IVIS Spectrum. This result showed that the bioluminescence signal of glioma in mice with the injection of 10 mg/ml DFP was significantly reduced (Figure 6E and F). The above results suggested that iron deficiency induced by DFP could suppress glioma in mice effectively. Furthermore, we assessed the survival time of the tumor-bearing mice with DFP or saline injection and found that 10 mg/ml DFP injection could significantly prolong the survival time of tumor-bearing mice (Figure 6G). Altogether, blocking iron into tumor is a potential therapy for tumor suppression.

Glial ferritin level is regulated by NBs

The aforementioned results have highlighted the significant role of glial ferritin in controlling NB number and proliferation. We next wonder whether NBs could provide feedback to regulate glial ferritin level. To answer this question, we analyzed ferritin level after altering NB proliferation. Previous studies have corroborated that NB proliferation was diminished by blocking Tor signaling pathway or enhanced by knocking down tumor suppressor genes, such as brat or numb (Bowman et al., 2008; Reichardt et al., 2018; Sousa-Nunes et al., 2011). When NBs were in low proliferative status, Fer2LCH subunits in the brain were significantly decreased, while Fer1HCH subunits declined slightly with no significance (Figure 7A-D). Conversely, highly proliferative NBs, able to form tumors, induced significantly increased level of the ferritin (Figure 7A-D). In summary, the production of ferritin in glial cells depends on the level of NB proliferation.

The level of ferritin is controlled by NB proliferation

(A) Fer1HCH subunit level in Drosophila CNS after manipulating proliferation.

(B) Fer2LCH subunit level in Drosophila CNS after manipulating proliferation.

(C) Quantification of Fer1HCH level in (A).

(D) Quantification of Fer2LCH level in (B).

(E) The model of glial ferritin regulating NBs.

(C and D, n=3; ns, not significant; *: p<0.05; **: p<0.01; ***: p<0.001)

Discussion

Here we demonstrate that glia regulates iron metabolism in the brain to sustain NBs and their proliferation. Glial ferritin coordinates Zip13 to provide iron for NBs. Once ferritin in glia is defective, NBs proliferate much slower and differentiate prematurely, implying that iron is a necessary metallic element for NBs. Iron participates in the synthesis of Fe-S clusters which activate aconitase required for TCA and facilitate electron-transfer processes involved in OxPhos. Since TCA and OxPhos are essential steps of ATP production, glial ferritin supplies iron into NBs to ensure energy production for maintaining NBs and their proliferation (Figure 7E).

NSCs and glial niche

Previous literatures have reported how glial niche regulates NB reactivation, proliferation, and survival. Blood-brain barrier (BBB) glia secretes dILP2 and dILP6 that bind on InR of NBs, which switches on Insulin signaling to reactivate NBs (Chell & Brand, 2010; Speder & Brand, 2014). In addition, Pvr (PDGFR ortholog) signaling in cortex glia sustains NB survival and proliferation in a PI3K-DE-cadherin-dependent manner (Read, 2018). However, it is unknown whether glial niche is required for maintaining the self-renewal of NBs. In this study, we demonstrate that glial niche maintains the self-renewal and proliferation of NBs through supplying the necessary microelement iron into NBs for enough energy production, which implies that mammalian glia likewise provides critical microelement for maintaining NSC function. Oligodendrocytes have been reported to secret ferritin to protect neuron (Mukherjee et al., 2020), however, it remains unknown whether the ferritin secretion is employed by oligodendrocytes to regulate NSCs. Astrocytes regulate NSC proliferation via secreting growth factors (Shetty, Hattiangady, & Shetty, 2005), but it is unclear whether ferritin from astrocytes regulates NSC proliferation and other NSC behaviors, such as self-renewal and differentiation. Based on our findings, it is feasible to regulate the self-renewal and differentiation of NSCs indirectly through niche cells, which would help to reduce the risk of directly manipulating NSCs to treat neural diseases.

Ferritin and iron

Our results show that ferritin generated outside the brain was not detected in the CNS and vice versa (data not shown), suggesting ferritin can hardly cross the BBB glia, possibly because ferritin is too large to permeate. In this way, BBB glia separates the labile systemic iron from local iron inside the brain to guarantee the normal brain development, as the neural cells, including NBs, require a stable iron level for continuous energy generation during neurogenesis.

Although the dominant function of ferritin is iron storage in mammal, increasing evidence suggests that ferritin also can transport iron. H-ferritin expression in astrocytes is necessary for proper oligodendrocyte development and myelination (Cheli et al., 2021), which implies H-ferritin probably transport iron from astrocytes to oligodendrocytes for myelination. In mice, ferritin heavy chain is secreted by oligodendrocytes via extracellular vesicles (Mukherjee et al., 2020), which possibly endows ferritin with the function of iron transport. Consistently, previous results showed that ferritin can deliver iron across the cultured BBB (Chiou et al., 2019; Fisher et al., 2007). Our results emphasize the critical role of ferritin in iron transport in the CNS. Future studies would be interesting to investigate whether ferritin is able to transport iron in the mammalian brain.

Ferritin specifically targets on the tumor cells, which has been applied to tumor visualization and targeted drug delivery (Fan et al., 2012; Fan et al., 2018). Our study shows that the block of ferritin or iron can inhibit tumor effectively, which offers an explanation that ferritin targets on tumor is to transport iron for maintaining the cell proliferation, highlighting a potential target for tumor suppression. In consideration of the role of ferritin in normal tissue, such as antioxidant defense function for neurons (Mukherjee et al., 2020), entire block of ferritin for suppressing tumor probably bring neurological disorder. Therefore, we should target on ferritin specifically in tumor to suppress tumor growth.

Iron is an essential component for Fe-S clusters which participate in TCA and OxPhos, providing energy required for self-renewal and proliferation of NBs. Previous study claimed that OxPhos is required for NB proliferation (van den Ameele & Brand, 2019). However, another research observed that OxPhos is dispensable in type II NBs (Bonnay et al., 2020). Our data imply that energy deficiency in NBs caused by iron shortage leads to NB loss and low proliferation possibly through simultaneous disruption of TCA and OxPhos. Consistently, iron-loaded ferritin nanoparticles can be used to improve the self-renewal ability and differentiation potential of human NSCs in vitro (Lee, Yang, Cho, & Cho, 2018). Additionally, previous study has shown that iron-loaded ferritin can promote the proliferation of cultured cell from Drosophila (Li, 2010). Furthermore, iron depletion was reported to suppress human cancer cell growth at G1/S by inducing cyclin D1 proteolysis (Nurtjahja-Tjendraputra, Fu, Phang, & Richardson, 2007). These studies suggest that iron, as a necessary microelement for energy production, is essential to NSC maintenance and proliferation.

Bicellular iron homeostasis

Iron homeostasis can be regulated systemically by the liver-derived hormone hepcidin and the iron exporter ferroportin, and it is also regulated cellularly by the IRP/IRE system (Hentze, Muckenthaler, Galy, & Camaschella, 2010). Here we propose a bicellular iron homeostasis between NB and glial niche cell, which is distinct from classic cellular and systemic iron homeostasis. The bicellular iron homeostasis requires an elaborate mechanism to guarantee normal neurogenesis.

In a proliferative state, NBs receive iron from glial cells via ferritin to fulfill the energy requirements for proliferation and self-renewal. Therefore, the iron required by NBs is mainly stored in the ferritin of glial cells and allocated as needed. The bicellular iron homeostasis mechanism makes glial cells the main pool for ferritin production, relieving NBs from the need to produce ferritin and transferring the potential risk of iron overload-induced oxidative stress to glial cells. Furthermore, glial cells can subsequently clear oxidative stress through lipid droplets (Bailey et al., 2015), ensuring normal NB function during neural development. It is intriguing to explore whether bicellular iron homeostasis exists in other stem cells and their niches.

When normal NBs are converted into immortal tumor NBs, becoming highly proliferative, they need more nutrients including iron (Bonnay et al., 2020; van den Ameele & Brand, 2019). In response, more ferritin is generated to transport iron from the glial cells to NBs. However, it is unclear how NBs communicate with glial cells to determine the level of ferritin production. It’s interesting to identify the specific cues from NBs that regulate the level of ferritin in future studies. In addition, iron distribution in Drosophila brain is not observed due to low iron concentration and technical limitations in iron detection, which results in a lack of direct evidence for iron trafficking via ferritin.